### 2017-04-18 paper.tex
Creating skeleton for paper

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% enumerate package lets us use letters instead of numbers
\usepackage{enumerate}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

\usepackage{color}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below
\newcommand{\fixme}[2]{\textsc{\textbf{{#1}: {#2}}}}
\newcommand{\recommend}[1]{\textit{#1}}
\newcommand{\withurl}[2]{{#1}\footnote{\texttt{#2}}}
\newcommand{\rulemajor}[1]{\section{#1}}
\newcommand{\ruleminor}[1]{\textbf{#1}}
\begin{document}
\vspace*{0.2in}

\begin{flushleft}
{\Large
\textbf\newline{Ten Simple Rules for Making Research Software More Robust}
}
\newline
\\
{Greg~Wilson}\textsuperscript{1, *},
\\
\textbf{1} Rangle.io / gvwilson@third-bit.com
\\
* Corresponding author.
\end{flushleft}

\section*{Abstract}

FIXME: abstract

\section*{Author Summary}

FIXME: author summary

\section*{Introduction}

FIXME: introduction

Cite \cite{steinmacher2015}

\rulemajor{Rule 1: FIXME}

FIXME \ruleminor{specific content}

\section*{Conclusion}

FIXME

\bibliography{paper}

\end{document}

### 2017-05-29 paper.tex
Moving body of paper into separate file
Converting to LaTeX

@@
 \newcommand{\rulemajor}[1]{\section{#1}}
-\newcommand{\ruleminor}[1]{\textbf{#1}}
 \begin{document}
@@
 {\Large
-\textbf\newline{Ten Simple Rules for Making Research Software More Robust}
+\textbf\newline{Ten Simple Rules for Collaborative Lesson Development}
 }
@@
 \\
-{Greg~Wilson}\textsuperscript{1, *},
+{Gabriel~A.~Devenyi}\textsuperscript{1},
+{R\'{e}mi~Emonet}\textsuperscript{2},
+{Rayna~M.~Harris}\textsuperscript{3},
+{Kate~Hertweck}\textsuperscript{4},
+{Damien~Irving}\textsuperscript{5},
+{Ian~Milligan}\textsuperscript{6},
+{Greg~Wilson}\textsuperscript{7, *}
 \\
-\textbf{1} Rangle.io / gvwilson@third-bit.com
+\textbf{1} Douglas Mental Health University Institute / FIXME \\
+\textbf{2} FIXME: affiliation and email \\
+\textbf{3} University of Texas at Austin / rayna.harris@utexas.edu \\
+\textbf{4} University of Texas at Tyler / khertweck@uttyler.edu \\
+\textbf{5} CSIRO / FIXME \\
+\textbf{6} University of Waterloo / FIXME email \\
+\textbf{7} Rangle.io / gvwilson@third-bit.com
 \\
@@
 
-FIXME: introduction
+Thousands of university lecturers teach subjects ranging from first year biology,
+to graduate-level courses in Indian film.
+Some might use a common textbook written by a single author or two,
+but other than that they usually develop and improve their course materials in isolation.
+It is staggering to think how many wheels are being re-invented,
+how the progress of teaching in these fields has been held back by
+the lack of community collaboration and sharing,
+and how much valuable time has been wasted.
 
-Cite \cite{steinmacher2015}
+This problem extends beyond university teaching,
+but it is curious that it is endemic in tertiary education,
+given that research depends so critically on collaboration and sharing,
+and that most researchers complain about how much time teaching takes away from research.
 
-\rulemajor{Rule 1: FIXME}
+The software community's model of collaborative code development is an alternative.
+The authors first encountered it in the context of
+lessons on basic lab skills for research computing
+in the sciences and humanities...
 
-FIXME \ruleminor{specific content}
+FIXME: fill in introduction
+
+\rulemajor{Clarify your audience.}
+
+The first requirement for building lessons together is
+to know who they are being built for.
+``Archaeology students'' is far too vague:
+are you and your collaborators thinking of
+first-year students who need an introduction to the field,
+graduate students who intend to specialize in the sub-discipline which is the lesson's focus,
+or someone in betwee?
+Prerequisite knowledge,
+equipment or software required,
+how much time learners will actually have:
+if different contributors believe different things about these,
+they will find it difficult or impossible to work together.
+
+Thinking systematically about difficulty levels can help manage expectations.
+For example,
+the Programming Historian project labels lessons ``beginner''",
+``intermediate'',
+and ``advanced''
+to help authors write at appropriate levels.
+However,
+this only works if there is prior agreement on what those terms mean.
+
+Rather than itemizing prior knowledge and learning objectives,
+it can be helpful to write \emph{learner profiles}.
+This technique is borrowed from user interface design
+to help authors think about their audience's needs
+and give them a shorthand for talking about specific cases.
+
+Learner profiles have five parts:
+the learner's general background,
+what they already know,
+what *they* think they want to do,
+how the material will help them,
+and any special needs they might have.
+A learner profile for a weekend programming workshop for new college students might be:
+
+\begin{enumerate}
+
+\item
+  Jorge has just moved from Costa Rica to Canada
+  to study agricultural engineering.
+
+\item
+  Other than using Excel, Word, and the Internet,
+  Jorge's most significant previous experience with computers is
+  helping his sister build a WordPress site for the family business.
+
+\item
+  Jorge needs to measure properties of soil from nearby farms
+  using a handheld device that sends text files to his computer.
+  Right now, Jorge has to open each file in Excel,
+  crop the first and last points,
+  and calculate an average.
+
+\item
+  This workshop will show Jorge how to write a little Python program
+  to read the data,
+  select the right values from each file,
+  and calculate the required statistics.
+
+\item
+  Jorge can read English proficiently,
+  but still struggles sometimes to keep up with spoken conversation
+  (especially if it involves a lot of new jargon).
+
+\end{enumerate}
+
+\rulemajor{Build community around lessons.}
+
+Lessons don't maintain themelves. 
+Unlike traditional academic work, 
+which may have shelf time measured 
+in years (if not longer), 
+technical lessons do not maintain themselves. 
+Versions change, dependencies break, 
+and what was cutting edge in 2017 may be dated 
+and less useful in 2018. 
+This needs to be clear to contributors, 
+so they understand that this lesson is just 
+the starting point. 
+Sustainability needs to be front of mind.
+
+The focus accordingly needs to be on creating a community. 
+Authors cannot be expected to maintain continual vigilence on a lesson,
+but this is necessary if one expectes continual use of a lesson! 
+By working online, creating opportunities for collaboration 
+and contribution, many eyes can keep the lesson usable. 
+As we note below, 
+this only works if you make space for 
+little contributions just as you do for larger ones.
+
+\rulemajor{Reduce, re-use, recycle.}
+
+Don't create a new lesson if there is an existing one that you can use or contribute to.
+Just as a scholar would not write a paper without a literature review,
+the same holds for lessons.
+Before writing that introduction to the Bash command line,
+for example,
+do a search:
+has anybody else written it?
+Is it complementary to your goals?
+Could it be tweaked or modified to meet your own goals?
+Could your planned lesson be tweaked to compliment the existing lessons so that topics aren't duplicated?
+
+FIXME: This is a good place to talk about discoverability as well as licensing
+
+Before re-using content,
+make sure to check that the lesson is licensed in an open manner.
+Both Programming Historian and Software Carpentry license their material under
+the Creative Commons - Attribution license
+(see https://creativecommons.org/licenses/by/2.0/).
+This allows people to share and adapt material for any purpose,
+even commercial ones,
+without asking permissions as long as they continue to share the content in a similar manner.
+
+The same questions of re-use come when thinking about recycling content of a lesson,
+such as images, data, figure, or code.
+Does the license cover that as well?
+If not,
+then ask permission,
+just as you would for any other material.
+
+As you collaboratively develop your own lessons,
+make sure to think about the collaborative contexts of the entire field.
+
+\rulemajor{Build modular lessons that can be re-purposed.}
+
+Every instructor's needs are different,
+so build small chunks that can be re-purposed in many ways. 
+A university lecturer in meteorology, for instance,
+might construct a course by bringing together lessons on differential equations,
+fluid mechanics and absorption spectroscopy. 
+This task is greatly simplified if existing courses on mathematics, 
+physics and chemistry consist of numerous small, discrete lessons,
+as opposed to a few large, monolithic lessons.
+
+Note that doing this shifts the instructor's burden from writing to finding and synthesizing.
+Both of which are easier if lessons have been designed by people with a shared world-view (rule 6),
+and if lessons clearly signal what they cover (Rule~1).
+In particular,
+if lessons reference specific points in the model curriculum guidelines promulgated by many professional societies,
+it can be much easier for people to find them and integrate them.
+
+Note also that individual lesson topics hinge on the learners having the proper prerequisite background,
+so this rule further emphasizes the importance of rule 1.
+
+\rulemajor{Encourage and empower contributors.}
+
+It takes a village to ensure
+the steady improvement and adaptation that will give a lesson a long life.
+Ensuring this, however, is not straightforward.
+
+Making the process explicit is key.
+New contributors require a straightforward and transparent introduction
+to understand the process of tweaking and adapting lessons.
+Licensing, code of conduct, governance, and contribution procedure
+all need to be explicit rather than implicit
+to lower the social barriers to contribution.
+
+Tools can help,
+especially if they support pre-merge review.
+Yet some of these tools come with a considerable up-front learning curve.
+GitHub with pull requests may be ideal,
+but allowing people to edit a Google Doc,
+or to facilitate Wiki-based editing,
+can help get conversations started.
+In a pinch,
+work with contributors to find a comfort zone:
+don't let a collaboration flounder on technological choice.
+But make sure not to overwhelm either:
+threaded discussion (or GitHub issues) will help increase the signal-to-noise ratio.
+
+Ultimately,
+community dynamics are more important than platform.
+We have seen the 90/9/1 rule:
+90\% of people watch, 9\% discuss, 1\% do.
+This requires a gentle on-ramp for new contributors,
+that we make it easy for people to submit errata and suggestions,
+and that editors may need to do triage to ensure that voices are heard.
+Finally, with so many voices and contributions come caution.
+Working in the open can be great,
+but can also unintentionally suppress voices.
+At Programming Historian,
+for example,
+an ombudsperson is available for private chats and facilitation.
+
+\rulemajor{Teach best practices for lesson development.}
+
+Decades of pedagogical research has yielded many insights into
+how best to build and deliver lessons \cite{hlw}.
+Unfortunately,
+since most college and university faculty have little or no training in education,
+this knowledge and expertise is rarely transfered into classroom practice.
+
+Experience shows that even a brief introduction to a handful of key practices
+can help collaborative lesson development in at least three ways.
+If people have a shared understanding of how lessons should be developed,
+it is easier for them to work together.
+Less obviously,
+if people have a shared model of how lessons are going to be *used*,
+they are more likely to try to build the same kind of material.
+Finally,
+teaching people how to teach is a great way to introduce them to each other and build community.
+
+An example of a particular lesson development practice is reverse instructional design
+\cite{wiggins-mctighe}.
+When this is used,
+lessons are built by:
+
+\begin{itemize}
+
+\item
+  identifying the learning objectives,
+
+\item
+  creating \emph{summative assessments} to determine whether those objectives have been met,
+
+\item
+  designing \emph{formative assessments} to gauge learners' progress
+  and give them a chance to practice key skills,
+
+\item
+  putting those formative assessments in order,
+  and
+
+\item
+  writing lessons to connect each to the next.
+
+\end{itemize}
+
+This methodology is effective in its own right,
+but the real benefit comes when everyone understands what is supposed to happen
+in what order.
+
+Another example is Software Carpentry's instructor training program.
+First offered in 2012,
+this gradually turned into a two-day course delivered both in-person and online
+\cite{lessons-learned}.
+This crash course was a natural complement to Software Carpentry's regular workshops:
+while the latter were created so that researchers wouldn't have to teach themselves how to program,
+the former ensures that they don't have to teach themselves how to teach.
+Instructor training now serves and an onboarding process to get everyone on the same page
+regarding who lessons are for,
+how they are delivered,
+and how they are maintained.
+
+\rulemajor{Publish periodically and recognize contributions.}
+
+Similar to publishing software,
+lessons should have releases of fixed content
+so that learners or instructors who may wish to use the material have a stable version to refer to
+for the duration of their use.
+These releases should be periodic
+so that improvements and adjustments are made available
+for new learners and instructors just starting their use.
+Periodic releases are also essential
+for enabling recognition of the contribution of authors and maintainers.
+
+The traditional academic system has limited ways of recognizing contributions.
+Until systems have been expanded to improve this,
+it is important to publish your lessons using a mechanism that provides recognition to the contributors.
+Currently an effective way to do this is
+to publish lessons with a DOI supplied by (FIXME list possibilities).
+Contributors can be listed as authors
+and the maintainers of the lesson as editors
+to provide differentiated recognition of their contributions.
+
+A lesson release is a good opportunity to bring the material into a stable shape
+by fixing outstanding issues and merging contribution.
+The complete name and possible identifier like ORCID should also be gathered for every contributor.
+Version control helps a lot in continuously maintaining a list of contributors
+but also in remembering which version is used for release
+(e.g., using branches or tags).
+A lesson release is here to stay
+and it is recommended to use a consistent naming scheme from the beginning.
+A convenient naming scheme is to use the full year and month of the release,
+like ``2017.05'',
+as it conveys a clear notion of date,
+even to people that are new to the project.
+
+Once the lesson is ready for release and has a name,
+it should be built
+and submitted to DOI supplier with its metadata,
+such as lesson name, authors (e.g., contributors) and editors (e.g., maintainers).
+Due to the collaborative aspect of the lesson development,
+the number of authors and even lessons can grow very fast
+and it is worth spending some time on automating this release process.
+After the formal publishing,
+it is also a good idea to publish the built lesson for direct access by the community,
+making it clear which version is displayed and what is its corresponding DOI.
+
+\rulemajor{Evaluate lessons at several scales.}
+
+The purpose of feedback is to guide lesson development
+so that authors aren't designing and arguing in a vacuum.
+What people immersed in the lessons think needs fixing
+can often differ from what learners think.
+
+Micro-scale feedback can be gathered by an instructor while teaching a particular lesson.
+Learners might provide feedback on things like typographical errors,
+clarity/ease of quiz questions and/or the order in which topics are presented,
+which the instructor can enter into a work-tracking system (e.g., GitHub issues) at the end of class.
+As well as encouraging direct verbal feedback,
+it's a good idea to provide learners with a means to provide feedback anonymously during class 
+(e.g., on small pieces of paper like sticky notes).
+
+Pre- and post-class surveys can be used to discover larger macro-scale issues.
+These issues often relate to the fact that it can be difficult for lesson developers
+to fully understand the frame of reference of their audience.
+For instance, a lesson might inadvertently assume prior knowledge that many learners don't have,
+which is information that can be collected easily in a survey.
+If possible, it's a good idea to conduct the post-class survey 30-60 days after the fact.
+This allows people time to reflect,
+meaning they are more likely to give accurate feedback on what they learned
+rather than how entertained they were.
+
+Pre- and post-class surveys also are essential in focusing in on the audience for a given lesson.
+It can sometimes be difficult for lesson developers to understand the frame of a given learner
+so surveys (particularly post-class surveys) can reveal hidden assumed knowledge
+that can be expanded on or acknowledged.
+
+\rulemajor{Point to alternate resources.}
+
+\begin{itemize}
+
+\item
+  FIXME: The lesson is a beginning, not an end, so point them at similar material and where to go next.
+
+\item
+  FIXME: Make it part of a much broader conversation.
+
+\end{itemize}
+
+\rulemajor{You can't please everyone.}
+
+No single lesson can be right for every learner.
+Two people with no prior knowledge of a specific subject
+may still be able to move at very different speeds
+because of different levels of general background knowledge.
+Similarly,
+lessons on ecology for learners in Utah and Florida
+will probably be more relatable if they use different examples.
+Equally,
+no lesson development community can serve all purposes.
+Some groups may prioritize rapid evolution,
+while others may prefer a ``measure twice, cut once'' approach.
+
+If there are several complementary ways to explain something,
+or several points of views that can cohabit respectfully,
+it may be possible to present them side by side.
+There are actualy good pedagogical reasons to do this even if contributors *don't* disagree:
+weighing alternatives fosters higher-order thinking,
+and as Caulfield has written,
+the most popular online question and answer sites
+are successful in part because they present a chorus of explanations
+geared at different levels and needs
+rather than a single one (FIXME: cite choral explanations).
+
+But sometimes choices must be made.
+The open source software community has wrestled with these issues for three decades,
+and has evolved some best practices to handle them
+\cite{producing-oss}.
+As discussed in Rule~5,
+the first step is to have a clear governance structure and a clear, permissive license.
+Minor disagreements should be discussed openly and respectfully.
+If they cannot be resolved---i.e., if they turn out not to be so minor after all---then
+contributors should split off and evolve the lesson in the way they see best.
+(This is one of the reasons to have a permissive license.)
+
+Forking rarely happens in practice.
+When it does,
+it is important to remember that we all share the same vision of better lessons, built together.
 

### 2017-05-29 paper.tex
Fixing typos

@@
 graduate students who intend to specialize in the sub-discipline which is the lesson's focus,
-or someone in betwee?
+or someone in between?
 Prerequisite knowledge,
@@
 
-Lessons don't maintain themelves. 
+Lessons don't maintain themselves. 
 Unlike traditional academic work, 
@@
 The focus accordingly needs to be on creating a community. 
-Authors cannot be expected to maintain continual vigilence on a lesson,
-but this is necessary if one expectes continual use of a lesson! 
+Authors cannot be expected to maintain continual vigilance on a lesson,
+but this is necessary if one expects continual use of a lesson! 
 By working online, creating opportunities for collaboration 
@@
 since most college and university faculty have little or no training in education,
-this knowledge and expertise is rarely transfered into classroom practice.
+this knowledge and expertise is rarely transferred into classroom practice.
 
@@
 it may be possible to present them side by side.
-There are actualy good pedagogical reasons to do this even if contributors *don't* disagree:
+There are actually good pedagogical reasons to do this even if contributors *don't* disagree:
 weighing alternatives fosters higher-order thinking,

### 2017-05-29 paper.tex
Affiliations
Updated Damien's affiliation

@@
 \\
-\textbf{1} Douglas Mental Health University Institute / FIXME \\
-\textbf{2} FIXME: affiliation and email \\
+\textbf{1} Douglas Mental Health University Institute / gdevenyi@gmail.com \\
+\textbf{2} Universit\'{e} Jean Monnet / remi.emonet@univ-st-etienne.fr \\
 \textbf{3} University of Texas at Austin / rayna.harris@utexas.edu \\
 \textbf{4} University of Texas at Tyler / khertweck@uttyler.edu \\
-\textbf{5} CSIRO / FIXME \\
-\textbf{6} University of Waterloo / FIXME email \\
+\textbf{5} CSIRO Oceans and Atmosphere / irving.damien@gmail.com \\
+\textbf{6} University of Waterloo / i2millig@uwaterloo.ca \\
 \textbf{7} Rangle.io / gvwilson@third-bit.com

### 2017-05-30 paper.tex
Add DOI providers, tweak wording

@@
 Currently an effective way to do this is
-to publish lessons with a DOI supplied by (FIXME list possibilities).
+to publish lessons with a DOI supplied by providers such as Zenodo\footnote{https://zenodo.org/} or DataCite\footnote{https://zenodo.org/}.
 Contributors can be listed as authors
@@
 (e.g., using branches or tags).
-A lesson release is here to stay
+A lesson release is a permanent snapshot of the lesson
 and it is recommended to use a consistent naming scheme from the beginning.

### 2017-05-30 paper.tex
Clarify feedback for target audience

--- 2017-05-29 paper.tex
@@
 Pre- and post-class surveys also are essential in focusing in on the audience for a given lesson.
-It can sometimes be difficult for lesson developers to understand the frame of a given learner
+Refering back to Rule~1, it can sometimes be difficult for lesson developers to understand the frame of a given learner
 so surveys (particularly post-class surveys) can reveal hidden assumed knowledge
-that can be expanded on or acknowledged.
+that can be expanded on or acknowledged to refine the target audience.
 

### 2017-05-30 paper.tex
Draft of Rule 9

--- 2017-05-29 paper.tex
@@
 
+Most learners are unlikely to completely absorb the topic you are trying to
+present using only your lesson.
+There are many angles and approaches for any given topic.
+Make an attempt to find these alternative resources and offer them at strategic points in the lesson,
+and collect them at the end.
+Some learners prefer different material styles, so seek out textbooks, technical
+documentation, videos and web pages which may present the material in alternative styles.
+If a community or discussion forum exists for the topic, be sure to include these resources.
+
+Beyond the topic of your lesson, learners can go in many directions after completion.
+Provider learners with extensions and expansions of the topic from a variety of
+applications and viewpoints.
+
 \begin{itemize}
@@
 \item
-  FIXME: The lesson is a beginning, not an end, so point them at similar material and where to go next.
+  The lesson is a beginning, not an end, for the learner
 
 \item
-  FIXME: Make it part of a much broader conversation.
+  point them at similar material
+
+\item
+  and where to go next when they want to expand their knowledge.
+
+\item
+  A lesson is one piece of a larger engagement your learners are having
+  with a topic. Enable their further engagement.
 

### 2017-05-30 paper.tex
Fix emphasis

--- 2017-05-30 paper.tex
@@
 what they already know,
-what *they* think they want to do,
+what \emph{they} think they want to do,
 how the material will help them,
@@
 Less obviously,
-if people have a shared model of how lessons are going to be *used*,
+if people have a shared model of how lessons are going to be \emph{used},
 they are more likely to try to build the same kind of material.
@@
 Pre- and post-class surveys also are essential in focusing in on the audience for a given lesson.
-Refering back to Rule~1, it can sometimes be difficult for lesson developers to understand the frame of a given learner
+It can sometimes be difficult for lesson developers to understand the frame of a given learner
 so surveys (particularly post-class surveys) can reveal hidden assumed knowledge
-that can be expanded on or acknowledged to refine the target audience.
+that can be expanded on or acknowledged.
 
@@
 it may be possible to present them side by side.
-There are actually good pedagogical reasons to do this even if contributors *don't* disagree:
+There are actually good pedagogical reasons to do this even if contributors \emph{don't} disagree:
 weighing alternatives fosters higher-order thinking,

### 2017-05-30 paper.tex
Fix rule/Rule formatting

@@
 Note that doing this shifts the instructor's burden from writing to finding and synthesizing.
-Both of which are easier if lessons have been designed by people with a shared world-view (rule 6),
+Both of which are easier if lessons have been designed by people with a shared world-view (Rule~6),
 and if lessons clearly signal what they cover (Rule~1).
@@
 Note also that individual lesson topics hinge on the learners having the proper prerequisite background,
-so this rule further emphasizes the importance of rule 1.
+so this rule further emphasizes the importance of Rule~1.
 

### 2017-05-30 paper.tex
Fix latex typography for captial at end of sentence

@@
 it is also a good idea to publish the built lesson for direct access by the community,
-making it clear which version is displayed and what is its corresponding DOI.
+making it clear which version is displayed and what is its corresponding DOI@.
 

### 2017-05-30 paper.tex
Fix URL typography

@@
 Both Programming Historian and Software Carpentry license their material under
-the Creative Commons - Attribution license
-(see https://creativecommons.org/licenses/by/2.0/).
+the Creative Commons - Attribution license\footnote{\url{https://creativecommons.org/licenses/by/2.0/}}.
 This allows people to share and adapt material for any purpose,

### 2017-05-30 paper.tex
Fix starting sentence with however
little to small for clearer meaning

@@
 to help authors write at appropriate levels.
-However,
-this only works if there is prior agreement on what those terms mean.
+This method works, however, only if there is prior agreement on what those terms mean.
 
@@
 \item
-  This workshop will show Jorge how to write a little Python program
+  This workshop will show Jorge how to write a small Python program
   to read the data,

### 2017-05-30 paper.tex
updated affiliation

--- 2017-05-30 paper.tex
@@
 \textbf{2} Universit\'{e} Jean Monnet / remi.emonet@univ-st-etienne.fr \\
-\textbf{3} University of Texas at Austin / rayna.harris@utexas.edu \\
-\textbf{4} University of Texas at Tyler / khertweck@uttyler.edu \\
+\textbf{3} The University of Texas at Austin / rayna.harris@utexas.edu \\
+\textbf{4} The University of Texas at Tyler / khertweck@uttyler.edu \\
 \textbf{5} CSIRO Oceans and Atmosphere / irving.damien@gmail.com \\
@@
 what they already know,
-what \emph{they} think they want to do,
+what *they* think they want to do,
 how the material will help them,
@@
 Note that doing this shifts the instructor's burden from writing to finding and synthesizing.
-Both of which are easier if lessons have been designed by people with a shared world-view (Rule~6),
+Both of which are easier if lessons have been designed by people with a shared world-view (rule 6),
 and if lessons clearly signal what they cover (Rule~1).
@@
 Note also that individual lesson topics hinge on the learners having the proper prerequisite background,
-so this rule further emphasizes the importance of Rule~1.
+so this rule further emphasizes the importance of rule 1.
 
@@
 Less obviously,
-if people have a shared model of how lessons are going to be \emph{used},
+if people have a shared model of how lessons are going to be *used*,
 they are more likely to try to build the same kind of material.
@@
 it is also a good idea to publish the built lesson for direct access by the community,
-making it clear which version is displayed and what is its corresponding DOI@.
+making it clear which version is displayed and what is its corresponding DOI.
 
@@
 it may be possible to present them side by side.
-There are actually good pedagogical reasons to do this even if contributors \emph{don't} disagree:
+There are actually good pedagogical reasons to do this even if contributors *don't* disagree:
 weighing alternatives fosters higher-order thinking,

### 2017-05-30 paper.tex
Update gdevenyi affiliation

@@
 \\
-\textbf{1} Douglas Mental Health University Institute / gdevenyi@gmail.com \\
+\textbf{1} Douglas Mental Health University Institute, McGill University / gdevenyi@gmail.com \\
 \textbf{2} Universit\'{e} Jean Monnet / remi.emonet@univ-st-etienne.fr \\
-\textbf{3} The University of Texas at Austin / rayna.harris@utexas.edu \\
-\textbf{4} The University of Texas at Tyler / khertweck@uttyler.edu \\
+\textbf{3} University of Texas at Austin / rayna.harris@utexas.edu \\
+\textbf{4} University of Texas at Tyler / khertweck@uttyler.edu \\
 \textbf{5} CSIRO Oceans and Atmosphere / irving.damien@gmail.com \\

### 2017-05-30 paper.tex

@@
 \\
-\textbf{1} Douglas Mental Health University Institute, McGill University / gdevenyi@gmail.com \\
+\textbf{1} Douglas Mental Health University Institute / gdevenyi@gmail.com \\
 \textbf{2} Universit\'{e} Jean Monnet / remi.emonet@univ-st-etienne.fr \\
@@
 Currently an effective way to do this is
-to publish lessons with a DOI supplied by (FIXME list possibilities).
+to publish lessons with a DOI supplied by providers such as Zenodo\footnote{\url{https://zenodo.org/}} or DataCite\footnote{\url{https://zenodo.org/}}.
 Contributors can be listed as authors

### 2017-05-30 paper.tex
Merging DOI provider links

@@
 %% Include all macros below
-\newcommand{\fixme}[2]{\textsc{\textbf{{#1}: {#2}}}}
-\newcommand{\recommend}[1]{\textit{#1}}
-\newcommand{\withurl}[2]{{#1}\footnote{\texttt{#2}}}
 \newcommand{\rulemajor}[1]{\section{#1}}
@@
 Currently an effective way to do this is
-to publish lessons with a DOI supplied by providers such as Zenodo\footnote{\url{https://zenodo.org/}} or DataCite\footnote{\url{https://zenodo.org/}}.
+to publish lessons with a DOI supplied by providers such as Zenodo (\url{https://zenodo.org/})
+or DataCite (\url{https://zenodo.org/}).
 Contributors can be listed as authors

### 2017-05-30 paper.tex

@@
 \\
-\textbf{1} Douglas Mental Health University Institute / gdevenyi@gmail.com \\
+\textbf{1} Douglas Mental Health University Institute, McGill University / gdevenyi@gmail.com \\
 \textbf{2} Universit\'{e} Jean Monnet / remi.emonet@univ-st-etienne.fr \\
-\textbf{3} University of Texas at Austin / rayna.harris@utexas.edu \\
-\textbf{4} University of Texas at Tyler / khertweck@uttyler.edu \\
+\textbf{3} The University of Texas at Austin / rayna.harris@utexas.edu \\
+\textbf{4} The University of Texas at Tyler / khertweck@uttyler.edu \\
 \textbf{5} CSIRO Oceans and Atmosphere / irving.damien@gmail.com \\

### 2017-05-30 paper.tex
Incorporating recent changes

--- 2017-05-30 paper.tex
@@
 %% Include all macros below
+\newcommand{\fixme}[2]{\textsc{\textbf{FIXME {#1}: {#2}}}}
 \newcommand{\rulemajor}[1]{\section{#1}}
@@
 
-FIXME: abstract
+\fixme{61}{Write abstract}
 
@@
 
-FIXME: author summary
+\fixme{62}{Write author summary}
 
@@
 
-FIXME: fill in introduction
+\fixme{63}{Complete the introduction}
 
@@
 
-FIXME: This is a good place to talk about discoverability as well as licensing
+\fixme{64}{Discuss discoverability as well as licensing}
 
@@
 Pre- and post-class surveys also are essential in focusing in on the audience for a given lesson.
-Refering back to Rule~1, it can sometimes be difficult for lesson developers to understand the frame of a given learner
+Refering back to Rule~1,
+it can sometimes be difficult for lesson developers to understand the frame of a given learner
 so surveys (particularly post-class surveys) can reveal hidden assumed knowledge
@@
 
-\rulemajor{Point to alternate resources.}
+\rulemajor{Link to other resources.}
 
-\begin{itemize}
+Most learners are unlikely to absorb everything there is to know about a topic
+from your lesson alone.
+This is partly a matter of scope---any interesting subject is too large
+to fit in a single lesson---but also a matter of level and direction.
+As Caulfield has argued \cite{choral-explanations},
+the best resources provide a chorus of explanations
+that offer many angles and approaches for any given topic,
+each of which may be the best fit for a different subset of people.
 
-\item
-  FIXME: The lesson is a beginning, not an end, so point them at similar material and where to go next.
+Find these resources and direct learners to them at strategic points.
+These resources may include textbooks,
+technical documentation,
+videos,
+and web pages;
+if a community or discussion forum exists for the topic,
+be sure to include them as well.
 
-\item
-  FIXME: Make it part of a much broader conversation.
-
-\end{itemize}
+Doing this is a lot of work,
+and maintaining it is even more,
+which makes building community around lessons (Rule~2) all the more important.
+In particular,
+it is vital to engage the learners as equal participants in that community:
+they should both be able to propose updates, corrections, and additions to lessons,
+and know that they are encouraged to do so (Rule~5).
 
@@
 geared at different levels and needs
-rather than a single one (FIXME: cite choral explanations).
+rather than a single one \cite{choral-explanations}.
 
@@
 
-FIXME
+\fixme{65}{Write conclusion}
 

### 2017-05-31 paper.tex
abstract #61 and author summary #62

@@
 
-\fixme{61}{Write abstract}
+Technical lessons development is hindered by several core problems. 
+Lessons can quickly become outdated, buggy, or fail to be updated. 
+Collaborative code development offers a pathway forward, allowing us to 
+create lessons which are both open and accessible, as well as 
+being continually maintained, refined, and improved by a large
+community of contributors. This paper provides ten simple rules that can help
+others create sustainable lessons.
 
@@
 
-\fixme{62}{Write author summary}
+The model of collaborative code development presents an 
+alternative model to traditional scientific lesson 
+development. By leveraging a community approach, educational 
+resources can be more sustainable, robust, and responsive. 
+These ten simple rules outline best practices for this model of 
+collaborative resource development.
 

### 2017-05-31 paper.tex
stab at #65

--- 2017-05-30 paper.tex
@@
 
-\fixme{65}{Write conclusion}
+Every day, university instructors and others are writing, teaching, sharing
+lessons on subjects as diverse as history, biology, mathematics, or 
+technical instruction. Our work overlaps, meaning thousands of hours of 
+unnecessary duplication. In these ten simple rules, we provide an alternative: 
+leveraging the model of collaborative software development to make more robust
+and sustainable lessons that can be continually improved. We hope
+that the lessons we provide here can help others learn from our experiences
+to help us all teach better.
 

### 2017-05-31 paper.tex
minor word change

@@
 unnecessary duplication. In these ten simple rules, we provide an alternative: 
-leveraging the model of collaborative software development to make more robust
+adopting the model of collaborative software development to make more robust
 and sustainable lessons that can be continually improved. We hope

### 2017-05-31 paper.tex
Update paper.tex

--- 2017-05-31 paper.tex
@@
 others create sustainable lessons. 
+Our ten rules focus on both the people and the mechanisms involved with 
+lesson development. First, we describe the importance of clarifying audiences, 
+building communities around lessons, teaching people how to develop lessons, 
+and empowering lesson contributors. Then we describe the utility of 
+building modular lessons, reusing lesson materials, pointing to external resources, 
+and publishing materials periodically. Finally, we encourage that all outcomes 
+are evaluated and assessed to ensure continues success of the program or 
+find new avenues for improvement.
 

### 2017-05-31 paper.tex

--- 2017-05-31 paper.tex
@@
 lessons on basic lab skills for research computing
-in the sciences and humanities...
+in the sciences and humanities. They primarily draw on 
+their experiences with pedagogical organizations such as 
+Software Carpentry and Programming Historian. 
+Software Carpentry was founded in 1998 with the goal of 
+``making scientists more productive, and their work more
+reliable, by teaching them basic computing skills.'' 
+They do so through short one or two day intensive workshops that
+cover the basics of scientific computing. Programming Historian 
+was founded in 2008, and has evolved to become a collaboratively edited
+site providing lessons to humanists.
+
+In this article, we strive for an open-source software model, 
+by which we mean lesson materials that are:
+
+\begin{enumerate}
+
+\item
+  Open and easily accessible
+
+\item
+  Continually maintained, refined, and 
+  improved by a community of contributors
+
+\end{enumerate}
+
+To our knowledge, Software Carpentry (and its related
+organizations such as Data Carpentry) are the only real-world
+example of an organization satisfying these criteria. Programming
+Historian, and other open educational resources, satisfy the
+first criteria but not necessarily the latter. While they 
+might experience a kind of iterative development where the 
+lessons are periodically updated by a small author group
+(e.g. this happens when a new edition of an academic 
+textbook is released), but this is not the same as regular, 
+small releases from a large community of contributors.
+
+Meeting both criteria is critical for two reasons. We need
+lessons that are up-to-date with the latest information,
+which is crucial given the speed of software development. A lag
+between editions means broken dependencies and bugs. Second, lessons
+need to continue to live after a key author moves on from 
+being able to maintain the lesson.
+
+This all means that the lessons have the potential to be far more useful.
+The collective knowledge and experience of the community
+is far greater than a single author. The ten simple rules 
+that follow aim to help others benefit from the
+lessons we have learned as maintainers, editors, and reviewers of 
+these resources.
 
@@
 
-Every day, university instructors and others are writing, teaching, sharing
-lessons on subjects as diverse as history, biology, mathematics, or 
-technical instruction. Our work overlaps, meaning thousands of hours of 
-unnecessary duplication. In these ten simple rules, we provide an alternative: 
-adopting the model of collaborative software development to make more robust
-and sustainable lessons that can be continually improved. We hope
-that the lessons we provide here can help others learn from our experiences
-to help us all teach better.
+\fixme{65}{Write conclusion}
 

### 2017-05-31 paper.tex
Merging and editing

@@
 Software Carpentry and Programming Historian.
-Software Carpentry was founded in 1998 with the goal of 
-``making scientists more productive, and their work more
-reliable, by teaching them basic computing skills.'' 
-They do so through short one or two day intensive workshops that
-cover the basics of scientific computing. Programming Historian 
-was founded in 2008, and has evolved to become a collaboratively edited
-site providing lessons to humanists.
+Software Carpentry was founded in 1998
+to make scientists more productive,
+and their work more reliable,
+by teaching them basic computing skills.
+It has since spawned two sibling organizations called Data Carpentry and Library Carpentry.
+Programming Historian was founded in 2008,
+and has evolved into a collaboratively-edited site providing lessons to humanists.
 
-In this article, we strive for an open-source software model, 
-by which we mean lesson materials that are:
+These organizations have been successful in adapting open source development methods
+to the problem of lesson development and maintenance.
+Their guiding principles are that lessons should be:
 
@@
 \item
-  Open and easily accessible
+  open and easily accessible, and
 
 \item
-  Continually maintained, refined, and 
-  improved by a community of contributors
+  continually maintained, refined, and improved
+  by a community of contributors.
 
@@
 
-To our knowledge, Software Carpentry (and its related
-organizations such as Data Carpentry) are the only real-world
-example of an organization satisfying these criteria. Programming
-Historian, and other open educational resources, satisfy the
-first criteria but not necessarily the latter. While they 
-might experience a kind of iterative development where the 
-lessons are periodically updated by a small author group
-(e.g. this happens when a new edition of an academic 
-textbook is released), but this is not the same as regular, 
-small releases from a large community of contributors.
+Many open education projects satisfy the first criterion,
+but very few satisfy the second.
+While their lessons might periodically be updated by a small author group,
+as happens when a new edition of a book is edited and published,
+this is not the same as regular, small releases
+from a large community of contributors.
 
-Meeting both criteria is critical for two reasons. We need
-lessons that are up-to-date with the latest information,
-which is crucial given the speed of software development. A lag
-between editions means broken dependencies and bugs. Second, lessons
-need to continue to live after a key author moves on from 
-being able to maintain the lesson.
+Meeting both criteria is critical for two reasons.
+We need lessons that are up-to-date with the latest information,
+which is crucial given the speed of software development.
+A lag between editions means broken dependencies and bugs.
+Second,
+lessons need to continue to live after a key author moves on
+from being able to maintain them.
 

### 2017-05-31 paper.tex

@@
 
-\fixme{61}{Write abstract}
+Technical lessons development is hindered by several core problems. 
+Lessons can quickly become outdated, buggy, or fail to be updated. 
+Collaborative code development offers a pathway forward, allowing us to 
+create lessons which are both open and accessible, as well as 
+being continually maintained, refined, and improved by a large
+community of contributors. This paper provides ten simple rules that can help
+others create sustainable lessons. 
+Our ten rules focus on both the people and the mechanisms involved with 
+lesson development. First, we describe the importance of clarifying audiences, 
+building communities around lessons, teaching people how to develop lessons, 
+and empowering lesson contributors. Then we describe the utility of 
+building modular lessons, reusing lesson materials, pointing to external resources, 
+and publishing materials periodically. Finally, we encourage that all outcomes 
+are evaluated and assessed to ensure continues success of the program or 
+find new avenues for improvement.
 

### 2017-05-31 paper.tex
Merging and editing abstract and author summary

--- 2017-05-31 paper.tex
@@
 
-\section*{Abstract}
+\begin{abstract}
 
-\fixme{61}{Write abstract}
+\noindent
+Lessons take significant effort to build,
+and even more effort to maintain.
+The collaborative code development methods pioneered by the open source community
+offer a way forward,
+allowing us to create lessons which are open,
+accessible,
+and conntinually updated and improved by a community of contributors.
+This paper provides ten simple rules that can help others create sustainable lessons.
+
+\end{abstract}
 

### 2017-05-31 paper.tex

@@
 
-\fixme{65}{Write conclusion}
+Every day, university instructors and others are writing, teaching, sharing
+lessons on subjects as diverse as history, biology, mathematics, or 
+technical instruction. Our work overlaps, meaning thousands of hours of 
+unnecessary duplication. In these ten simple rules, we provide an alternative: 
+adopting the model of collaborative software development to make more robust
+and sustainable lessons that can be continually improved. We hope
+that the lessons we provide here can help others learn from our experiences
+to help us all teach better.
 

### 2017-05-31 paper.tex
Editing

--- 2017-05-31 paper.tex
@@
 
-\fixme{65}{Write conclusion}
+Every day,
+teachers all over the world spend countless hours duplicating each other's work.
+These ten rules provide an alternative:
+adopting the model of collaborative software development
+to make more robust and sustainable lessons
+that can be continually improved by those who use them.
+We hope that our experiences can help others teach more
+with more impact and less effort.
 

### 2017-06-01 paper.tex
Light editing

@@
 %% Include all macros below
-\newcommand{\fixme}[2]{\textsc{\textbf{FIXME {#1}: {#2}}}}
 \newcommand{\rulemajor}[1]{\section{#1}}
@@
 accessible,
-and conntinually updated and improved by a community of contributors.
+and continually updated and improved by a community of contributors.
 This paper provides ten simple rules that can help others create sustainable lessons.
@@
 Software Carpentry and Programming Historian.
-Software Carpentry was founded in 1998
-to make scientists more productive,
-and their work more reliable,
-by teaching them basic computing skills.
-It has since spawned two sibling organizations called Data Carpentry and Library Carpentry.
+Software Carpentry was founded in 1998 to teach scientists basic computing skills,
+and has since spawned two sibling organizations called Data Carpentry and Library Carpentry.
 Programming Historian was founded in 2008,
@@
 These organizations have been successful in adapting open source development methods
-to the problem of lesson development and maintenance.
+to lesson development and maintenance.
 Their guiding principles are that lessons should be:
@@
 
-Many open education projects satisfy the first criterion,
-but very few satisfy the second.
-While their lessons might periodically be updated by a small author group,
+By definition,
+open education projects satisfy the first criterion,
+but very few satisfy the second:
+while their lessons might occasionally be updated by a small author group,
 as happens when a new edition of a book is edited and published,
-this is not the same as regular, small releases
-from a large community of contributors.
-
-Meeting both criteria is critical for two reasons.
-We need lessons that are up-to-date with the latest information,
-which is crucial given the speed of software development.
-A lag between editions means broken dependencies and bugs.
-Second,
-lessons need to continue to live after a key author moves on
-from being able to maintain them.
-
-This all means that the lessons have the potential to be far more useful.
-The collective knowledge and experience of the community
-is far greater than a single author. The ten simple rules
-that follow aim to help others benefit from the
-lessons we have learned as maintainers, editors, and reviewers of
-these resources.
-
-\fixme{63}{Complete the introduction}
+this is not the same as continuous improvement by a large community of contributors.
+The ten simple rules that follow summarize what we have learned about doing that
+as maintainers, editors, and reviewers of lessons used by tens of thousands of people.
 
@@
 
-\fixme{64}{Discuss discoverability as well as licensing}
-
 Before re-using content,
 make sure to check that the lesson is licensed in an open manner.
-Both Programming Historian and Software Carpentry license their material under
-the Creative Commons - Attribution license
-(see https://creativecommons.org/licenses/by/2.0/).
-This allows people to share and adapt material for any purpose,
-even commercial ones,
-without asking permissions as long as they continue to share the content in a similar manner.
+Both Programming Historian and the various Carpentry projects
+use the Creative Commons - Attribution license
+(see https://creativecommons.org/licenses/by/4.0/),
+which allows people to share and adapt material for any purpose
+as long as they cite the original source.
 
@@
 
-As you collaboratively develop your own lessons,
-make sure to think about the collaborative contexts of the entire field.
+The converse of this rule is to license your own lessons in a similar open manner,
+and to make them discoverable.
+For example,
+when lessons are published (Rule~7),
+make sure that have the usual keywords in their bibliographic entries
+and HTML page headers.
 
@@
 Pre- and post-class surveys also are essential in focusing in on the audience for a given lesson.
-Refering back to Rule~1,
+Referring back to Rule~1,
 it can sometimes be difficult for lesson developers to understand the frame of a given learner

### 2017-06-01 paper.tex
Removing some Markdownisms

@@
 it may be possible to present them side by side.
-There are actually good pedagogical reasons to do this even if contributors *don't* disagree:
+There are actually good pedagogical reasons to do this
+even if contributors \emph{don't} disagree:
 weighing alternatives fosters higher-order thinking,

### 2017-06-01 paper.tex
Regularizing name

@@
 For example,
-the Programming Historian project labels lessons ``beginner''",
+Programming Historian project labels lessons ``beginner''",
 ``intermediate'',

### 2017-06-01 paper.tex
Removing description of learner profiles

@@
 Rather than itemizing prior knowledge and learning objectives,
-it can be helpful to write \emph{learner profiles}.
-This technique is borrowed from user interface design
-to help authors think about their audience's needs
-and give them a shorthand for talking about specific cases.
-
-Learner profiles have five parts:
+it can be helpful to write \emph{learner profiles} to clarify
 the learner's general background,
@@
 and any special needs they might have.
-A learner profile for a weekend programming workshop for new college students might be:
-
-\begin{enumerate}
-
-\item
-  Jorge has just moved from Costa Rica to Canada
-  to study agricultural engineering.
-
-\item
-  Other than using Excel, Word, and the Internet,
-  Jorge's most significant previous experience with computers is
-  helping his sister build a WordPress site for the family business.
-
-\item
-  Jorge needs to measure properties of soil from nearby farms
-  using a handheld device that sends text files to his computer.
-  Right now, Jorge has to open each file in Excel,
-  crop the first and last points,
-  and calculate an average.
-
-\item
-  This workshop will show Jorge how to write a little Python program
-  to read the data,
-  select the right values from each file,
-  and calculate the required statistics.
-
-\item
-  Jorge can read English proficiently,
-  but still struggles sometimes to keep up with spoken conversation
-  (especially if it involves a lot of new jargon).
-
-\end{enumerate}
 

### 2017-06-01 paper.tex
Editing down

@@
 make sure to check that the lesson is licensed in an open manner.
-Both Programming Historian and the various Carpentry projects
+Both Programming Historian and the Carpentry projects
 use the Creative Commons - Attribution license
-(see https://creativecommons.org/licenses/by/4.0/),
+(\url{https://creativecommons.org/licenses/by/4.0/}),
 which allows people to share and adapt material for any purpose
@@
 
-An example of a particular lesson development practice is reverse instructional design
+An example of a particular lesson development practice is \emph{reverse instructional design}
 \cite{wiggins-mctighe}.
 When this is used,
-lessons are built by:
-
-\begin{itemize}
-
-\item
-  identifying the learning objectives,
-
-\item
+lessons are built by
+identifying learning objectives,
 creating \emph{summative assessments} to determine whether those objectives have been met,
-
-\item
 designing \emph{formative assessments} to gauge learners' progress
 and give them a chance to practice key skills,
-
-\item
 putting those formative assessments in order,
-  and
-
-\item
-  writing lessons to connect each to the next.
-
-\end{itemize}
-
-This methodology is effective in its own right,
-but the real benefit comes when everyone understands what is supposed to happen
-in what order.
+and only then writing lessons to connect each to the next.
+This method is effective in its own right,
+but its greatest benefit is that it gives everyone a common framework
+within which to collaborate.
 
@@
 First offered in 2012,
-this gradually turned into a two-day course delivered both in-person and online
+is now a two-day course delivered both in-person and online
 \cite{lessons-learned}.
-This crash course was a natural complement to Software Carpentry's regular workshops:
-while the latter were created so that researchers wouldn't have to teach themselves how to program,
-the former ensures that they don't have to teach themselves how to teach.
-Instructor training now serves and an onboarding process to get everyone on the same page
+Not only does it teach good pedagogical practices,
+it serves and an onboarding process to get everyone on the same page
 regarding who lessons are for,
@@
 
-Similar to publishing software,
+Like software,
 lessons should have releases of fixed content
@@
 
-The traditional academic system has limited ways of recognizing contributions.
-Until systems have been expanded to improve this,
-it is important to publish your lessons using a mechanism that provides recognition to the contributors.
-Currently an effective way to do this is
-to publish lessons with a DOI supplied by providers such as Zenodo (\url{https://zenodo.org/})
+Academia has only a few ways of recognizing contributions.
+Until these are expanded,
+it is important to publish lessons in ways that traditional academic systems can digest.
+One is to give particular releases of lessons
+DOIs supplied by providers such as Zenodo (\url{https://zenodo.org/})
 or DataCite (\url{https://zenodo.org/}).
@@
 and the maintainers of the lesson as editors
-to provide differentiated recognition of their contributions.
+to differentiate recognition of their contributions.
+Each time the lesson is published,
+names and (where possible) identifiers like ORCIDs
+should also be gathered for all contributors.
 
 A lesson release is a good opportunity to bring the material into a stable shape
-by fixing outstanding issues and merging contribution.
-The complete name and possible identifier like ORCID should also be gathered for every contributor.
+by fixing outstanding issues and merging contributions.
 Version control helps a lot in continuously maintaining a list of contributors
@@
 (e.g., using branches or tags).
-A lesson release is a permanent snapshot of the lesson
-and it is recommended to use a consistent naming scheme from the beginning.
-A convenient naming scheme is to use the full year and month of the release,
-like ``2017.05'',
-as it conveys a clear notion of date,
-even to people that are new to the project.
+Lesson releases should use a consistent naming scheme,
+such as the full year and month of the release,
+e.g. ``2017.05''.
 
-Once the lesson is ready for release and has a name,
-it should be built
-and submitted to DOI supplier with its metadata,
-such as lesson name, authors (e.g., contributors) and editors (e.g., maintainers).
-Due to the collaborative aspect of the lesson development,
-the number of authors and even lessons can grow very fast
-and it is worth spending some time on automating this release process.
-After the formal publishing,
-it is also a good idea to publish the built lesson for direct access by the community,
-making it clear which version is displayed and what is its corresponding DOI.
+If lessons are being released regularly,
+it is worth automating this release process
+with a simple shell script.
+Old versions of lessons should be archived in a discoverable location for reference.
 

### 2017-06-02 paper.tex
fixed DataCite url, added ORCID url

@@
 DOIs supplied by providers such as Zenodo (\url{https://zenodo.org/})
-or DataCite (\url{https://zenodo.org/}).
+or DataCite (\url{https://www.datacite.org}).
 Contributors can be listed as authors
@@
 Each time the lesson is published,
-names and (where possible) identifiers like ORCIDs
-should also be gathered for all contributors.
+names and identifiers (e.g. ORCIDs (\url{https://orcid.org}))
+should be gathered for all contributors.
 

### 2017-06-05 paper.tex
A series of minor fixes to remove typos and jargon

--- 2017-06-01 paper.tex
@@
 
-Thousands of university lecturers teach subjects ranging from first year biology,
+Thousands of university lecturers around the world teach subjects ranging from first year biology
 to graduate-level courses in Indian film.
 Some might use a common textbook written by a single author or two,
-but other than that they usually develop and improve their course materials in isolation.
-It is staggering to think how many wheels are being re-invented,
-how the progress of teaching in these fields has been held back by
-the lack of community collaboration and sharing,
-and how much valuable time has been wasted.
+but besides that lecturers usually develop and improve their course materials in isolation.
+It is staggering to think how many wheels have been re-invented,
+how much valuable time has been wasted,
+and ultimately how the progress of tertiary education has been held back by
+the lack of community collaboration and sharing in this process.
 
-This problem extends beyond university teaching,
+This problem obviously extends beyond the university sector,
 but it is curious that it is endemic in tertiary education,
@@
 By definition,
-open education projects satisfy the first criterion,
+many open education projects satisfy the first criterion,
 but very few satisfy the second:
@@
 
-Note that doing this shifts the instructor's burden from writing to finding and synthesizing.
-Both of which are easier if lessons have been designed by people with a shared world-view (Rule~6),
+Doing this shifts the instructor's burden from writing to finding and synthesizing,
+both of which are easier if lessons have been designed by people with a shared world-view (Rule~6),
 and if lessons clearly signal what they cover (Rule~1).
@@
 Tools can help,
-especially if they support pre-merge review.
+especially if they allow proposed changes to be viewed and discussed
+prior to their incorporation/merger into the lessons
+(in open source software development this is known as ``pre-merge review'').
 Yet some of these tools come with a considerable up-front learning curve.
-GitHub with pull requests may be ideal,
-but allowing people to edit a Google Doc,
-or to facilitate Wiki-based editing,
-can help get conversations started.
+GitHub with pull requests is ideal for pre-merge review, for instance,
+but it requires contributors to understand the basics of the git version control software.
+Allowing people to edit a Google Doc or Wiki-based editing does not offer pre-merge review,
+but the low barrier to entry can help get conversations started.
 In a pinch,
@@
 But make sure not to overwhelm either:
-threaded discussion (or GitHub issues) will help increase the signal-to-noise ratio.
+threaded discussion on paticular topics (e.g. GitHub issues) can help increase the signal-to-noise ratio
+by reducing long reply-all email excahnges.
 
@@
 DOIs supplied by providers such as Zenodo (\url{https://zenodo.org/})
-or DataCite (\url{https://zenodo.org/}).
+or DataCite (\url{https://www.datacite.org/}).
 Contributors can be listed as authors
@@
 For instance, a lesson might inadvertently assume prior knowledge that many learners don't have,
-which is information that can be collected easily in a survey.
-If possible, it's a good idea to conduct the post-class survey 30-60 days after the fact.
+which is information that can be collected in a survey.
+If possible, it's a good idea to conduct the post-class survey 30--60 days after the fact.
 This allows people time to reflect,
@@
 Referring back to Rule~1,
-it can sometimes be difficult for lesson developers to understand the frame of a given learner
+it can sometimes be difficult for lesson developers to understand the frame of a given learner,
 so surveys (particularly post-class surveys) can reveal hidden assumed knowledge
@@
 Doing this is a lot of work,
-and maintaining it is even more,
+and maintaining it is even more so,
 which makes building community around lessons (Rule~2) all the more important.

### 2017-06-05 paper.tex
incorporating Rayna's changes to the intro

@@
 
-Thousands of university lecturers around the world teach subjects ranging from first year biology
+Lessons take significant time and energy to build and even more effort to maintain.
+Collaborative lesson development refers to the combined efforts of teachers to build and maintain lessons.
+By leveraging a community approach, educational resources can be more sustainable, robust, and responsive. 
+When lessons are created in an open source community, all lessons are open and accessible, 
+and they can be continually updated and improved by a community of contibutors.
+
+Despite the proven success of open-source and collaboratively developed lesson plans, 
+it is surprisingly uncommon in the traditional accademic setting.
+Each year, thousands of university lecturers teach subjects ranging from first year biology,
 to graduate-level courses in Indian film.
 Some might use a common textbook written by a single author or two,
-but besides that lecturers usually develop and improve their course materials in isolation.
-It is staggering to think how many wheels have been re-invented,
+but besides that lecturers develop and improve their course materials in isolation.
+It is staggering to think how many wheels are being re-invented,
 how much valuable time has been wasted,
@@
 the lack of community collaboration and sharing in this process.
-
 This problem obviously extends beyond the university sector,
@@
 
-The software community's model of collaborative code development is an alternative.
-The authors first encountered it in the context of
-lessons on basic lab skills for research computing
-in the sciences and humanities. They primarily draw on
-their experiences with pedagogical organizations such as
+The authors collectively have experience with community-developed lesson plans
+in the context of research computing in the sciences and humanities. 
+They primarily draw on their experiences with pedagogical organizations such as
 Software Carpentry and Programming Historian.
@@
 Programming Historian was founded in 2008,
-and has evolved into a collaboratively-edited site providing lessons to humanists.
-
-These organizations have been successful in adapting open source development methods
+and has evolved into a collaboratively-edited site providing lessons to humanists. These organizations have been successful in adapting open source development methods
 to lesson development and maintenance.

### 2017-06-05 paper.tex

@@
 Lessons don't maintain themselves.
-Unlike traditional academic work,
-which may have shelf time measured
-in years (if not longer),
-technical lessons do not maintain themselves.
-Versions change, dependencies break,
-and what was cutting edge in 2017 may be dated
+For technical lessons software versions and dependencies are constantly changing,
+while for more traditional academic lessons the literature is advancing at an ever increasing pace.
+In either case, what was cutting edge in 2017 may be dated
 and less useful in 2018.

### 2017-06-07 paper.tex
Getting latest version of template

@@
+% Template for PLoS
+% Version 3.4 January 2017
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
 \documentclass[10pt,letterpaper]{article}
@@
 
-% enumerate package lets us use letters instead of numbers
-\usepackage{enumerate}
-
 % create "+" rule type for thick vertical lines
@@
 \noalign{\global\arrayrulewidth\savedwidth}}
-
-\usepackage{color}
 
@@
 %% Include all macros below
+
+% \newcommand{\lorem}{{\bf LOREM}}
+% \newcommand{\ipsum}{{\bf IPSUM}}
 \newcommand{\rulemajor}[1]{\section{#1}}
+
+%% END MACROS SECTION
+
 \begin{document}
@@
 
+% Title must be 250 characters or less.
 \begin{flushleft}
 {\Large
-\textbf\newline{Ten Simple Rules for Collaborative Lesson Development}
+\textbf\newline{Ten simple rules for collaborative lesson development} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
 \newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
 \\
-{Gabriel~A.~Devenyi}\textsuperscript{1},
-{R\'{e}mi~Emonet}\textsuperscript{2},
-{Rayna~M.~Harris}\textsuperscript{3},
-{Kate~Hertweck}\textsuperscript{4},
-{Damien~Irving}\textsuperscript{5},
-{Ian~Milligan}\textsuperscript{6},
-{Greg~Wilson}\textsuperscript{7, *}
+{Gabriel~A.~Devenyi}\textsuperscript{1{\ddag}},
+{R\'{e}mi~Emonet}\textsuperscript{2{\ddag}},
+{Rayna~M.~Harris}\textsuperscript{3{\ddag}},
+{Kate~Hertweck}\textsuperscript{4{\ddag}},
+{Damien~Irving}\textsuperscript{5{\ddag}},
+{Ian~Milligan}\textsuperscript{6{\ddag}},
+{Greg~Wilson}\textsuperscript{7{\ddag}*}
 \\
+\bigskip
 \textbf{1} Douglas Mental Health University Institute, McGill University / gdevenyi@gmail.com \\
@@
 \textbf{6} University of Waterloo / i2millig@uwaterloo.ca \\
-\textbf{7} Rangle.io / gvwilson@third-bit.com
-\\
-* Corresponding author.
+\textbf{7} Rangle.io / gvwilson@third-bit.com \\
+
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+{\ddag} These authors contributed equally to this work.
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+% \ddag These authors also contributed equally to this work.
+
+% Current address notes
+% \textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
+% \textcurrency b Insert second current address 
+% \textcurrency c Insert third current address
+
+% Deceased author note
+% \dag Deceased
+
+% Group/Consortium Author Note
+% \textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* corresponding author
+
 \end{flushleft}
 
-\begin{abstract}
+% Please keep the abstract below 300 words
 
-\noindent
+\linenumbers
+
+\section*{Abstract}
+
 Lessons take significant effort to build,
@@
 
-\end{abstract}
-
-\section*{Author Summary}
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+\section*{Author summary}
 
@@
 collaborative resource development.
+
+% Use "Eq" instead of "Equation" for equation citations.
 
@@
 
-\rulemajor{Clarify your audience.}
+\rulemajor{Clarify your audience}
 
@@
 
-\rulemajor{Build community around lessons.}
+\rulemajor{Build community around lessons}
 
@@
 
-\rulemajor{Reduce, re-use, recycle.}
+\rulemajor{Reduce, re-use, recycle}
 
@@
 
-\rulemajor{Build modular lessons that can be re-purposed.}
+\rulemajor{Build modular lessons that can be re-purposed}
 
@@
 
-\rulemajor{Encourage and empower contributors.}
+\rulemajor{Encourage and empower contributors}
 
@@
 
-\rulemajor{Teach best practices for lesson development.}
+\rulemajor{Teach best practices for lesson development}
 
@@
 
-\rulemajor{Publish periodically and recognize contributions.}
+\rulemajor{Publish periodically and recognize contributions}
 
@@
 
-\rulemajor{Evaluate lessons at several scales.}
+\rulemajor{Evaluate lessons at several scales}
 
@@
 
-\rulemajor{Link to other resources.}
+\rulemajor{Link to other resources}
 
@@
 
-\rulemajor{You can't please everyone.}
+\rulemajor{You can't please everyone}
 
@@
 
-\bibliography{paper}
+\nolinenumbers
+
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% or
+%
+% Compile your BiBTeX database using our plos2015.bst
+% style file and paste the contents of your .bbl file
+% here. See http://journals.plos.org/plosone/s/latex for 
+% step-by-step instructions.
+%
+
+\begin{thebibliography}{1}
+
+\bibitem{hlw}
+Ambrose SA, Bridges MW, DiPietro M, Lovett MC, Norman MK.
+\newblock How Learning Works.
+\newblock Jossey-Bass; 2010.
+
+\bibitem{wiggins-mctighe}
+Wiggins G, McTighe J.
+\newblock Understanding by Design.
+\newblock 2nd ed. Association for Supervision and Curriculum Development; 2005.
+
+\bibitem{lessons-learned}
+Wilson G.
+\newblock Software Carpentry: Lessons Learned.
+\newblock F1000Research. 2016;3(62).
+\newblock doi:{10.12688/f1000research.3-62.v2}.
+
+\bibitem{choral-explanations}
+Caulfield M. Choral Explanations; 2016.
+\newblock https://hapgood.us/2016/05/13/choral-explanations/.
+
+\bibitem{producing-oss}
+Fogel K.
+\newblock Producing Open Source Software.
+\newblock O'Reilly; 2005.
+
+\end{thebibliography}
 

### 2017-06-07 paper.tex
Updating to latest version of PLOS template

@@
 is now a two-day course delivered both in-person and online
-\cite{lessons-learned}.
+\cite{lessons-learned,instructor-training,how-to-teach-programming}.
 Not only does it teach good pedagogical practices,
@@
 e.g. ``2017.05''.
+This is the scheme Software Carpentry has used in
+its releases of its lessons \cite{shell2015,shell2017}.
 
@@
 % step-by-step instructions.
-%
 
@@
 
+\bibitem{instructor-training}
+Christina Koch and Greg Wilson (eds.)
+\newblock Software Carpentry: Instructor Training; 2016.
+\newblock \url{https://zenodo.org/record/57571#.WS8huDOZPdQ}.
+
+\bibitem{how-to-teach-programming}
+Wilson G.
+\newblock How to Teach Programming (And Other Things).
+\newblock Lulu.com; 2017.
+
+\bibitem{shell2015}
+Gabriel A.~Devenyi and Christina Koch (eds.):
+Software Carpentry: The Unix Shell; 2015.
+\newblock \url{https://zenodo.org/record/27355#.WS8lajOZPdQ}.
+
+\bibitem{shell2017}
+Gabriel A.~Devenyi and Ashwin Srinath (eds.):
+Software Carpentry: The Unix Shell; 2017.
+\newblock \url{https://zenodo.org/record/278226#.WS74tTOZPdQ}.
+
 \bibitem{choral-explanations}

### 2017-06-10 paper.tex
updates the rule order

@@
 
-\rulemajor{Reduce, re-use, recycle}
-
-Don't create a new lesson if there is an existing one that you can use or contribute to.
-Just as a scholar would not write a paper without a literature review,
-the same holds for lessons.
-Before writing that introduction to the Bash command line,
-for example,
-do a search:
-has anybody else written it?
-Is it complementary to your goals?
-Could it be tweaked or modified to meet your own goals?
-Could your planned lesson be tweaked to compliment the existing lessons so that topics aren't duplicated?
-
-Before re-using content,
-make sure to check that the lesson is licensed in an open manner.
-Both Programming Historian and the Carpentry projects
-use the Creative Commons - Attribution license
-(\url{https://creativecommons.org/licenses/by/4.0/}),
-which allows people to share and adapt material for any purpose
-as long as they cite the original source.
-
-The same questions of re-use come when thinking about recycling content of a lesson,
-such as images, data, figure, or code.
-Does the license cover that as well?
-If not,
-then ask permission,
-just as you would for any other material.
-
-The converse of this rule is to license your own lessons in a similar open manner,
-and to make them discoverable.
-For example,
-when lessons are published (Rule~7),
-make sure that have the usual keywords in their bibliographic entries
-and HTML page headers.
-
 \rulemajor{Build modular lessons that can be re-purposed}
@@
 Doing this shifts the instructor's burden from writing to finding and synthesizing,
-both of which are easier if lessons have been designed by people with a shared world-view (Rule~6),
+both of which are easier if lessons have been designed by people with a shared world-view (Rule~4),
 and if lessons clearly signal what they cover (Rule~1).
@@
 so this rule further emphasizes the importance of Rule~1.
+
+\rulemajor{Teach best practices for lesson development}
+
+Decades of pedagogical research has yielded many insights into
+how best to build and deliver lessons \cite{hlw}.
+Unfortunately,
+since most college and university faculty have little or no training in education,
+this knowledge and expertise is rarely transferred into classroom practice.
+
+Experience shows that even a brief introduction to a handful of key practices
+can help collaborative lesson development in at least three ways.
+If people have a shared understanding of how lessons should be developed,
+it is easier for them to work together.
+Less obviously,
+if people have a shared model of how lessons are going to be \emph{used},
+they are more likely to try to build the same kind of material.
+Finally,
+teaching people how to teach is a great way to introduce them to each other and build community.
+
+An example of a particular lesson development practice is \emph{reverse instructional design}
+\cite{wiggins-mctighe}.
+When this is used,
+lessons are built by
+identifying learning objectives,
+creating \emph{summative assessments} to determine whether those objectives have been met,
+designing \emph{formative assessments} to gauge learners' progress
+and give them a chance to practice key skills,
+putting those formative assessments in order,
+and only then writing lessons to connect each to the next.
+This method is effective in its own right,
+but its greatest benefit is that it gives everyone a common framework
+within which to collaborate.
+
+Another example is Software Carpentry's instructor training program.
+First offered in 2012,
+is now a two-day course delivered both in-person and online
+\cite{lessons-learned,instructor-training,how-to-teach-programming}.
+Not only does it teach good pedagogical practices,
+it serves and an onboarding process to get everyone on the same page
+regarding who lessons are for,
+how they are delivered,
+and how they are maintained.
 
@@
 an ombudsperson is available for private chats and facilitation.
-
-\rulemajor{Teach best practices for lesson development}
-
-Decades of pedagogical research has yielded many insights into
-how best to build and deliver lessons \cite{hlw}.
-Unfortunately,
-since most college and university faculty have little or no training in education,
-this knowledge and expertise is rarely transferred into classroom practice.
-
-Experience shows that even a brief introduction to a handful of key practices
-can help collaborative lesson development in at least three ways.
-If people have a shared understanding of how lessons should be developed,
-it is easier for them to work together.
-Less obviously,
-if people have a shared model of how lessons are going to be \emph{used},
-they are more likely to try to build the same kind of material.
-Finally,
-teaching people how to teach is a great way to introduce them to each other and build community.
-
-An example of a particular lesson development practice is \emph{reverse instructional design}
-\cite{wiggins-mctighe}.
-When this is used,
-lessons are built by
-identifying learning objectives,
-creating \emph{summative assessments} to determine whether those objectives have been met,
-designing \emph{formative assessments} to gauge learners' progress
-and give them a chance to practice key skills,
-putting those formative assessments in order,
-and only then writing lessons to connect each to the next.
-This method is effective in its own right,
-but its greatest benefit is that it gives everyone a common framework
-within which to collaborate.
-
-Another example is Software Carpentry's instructor training program.
-First offered in 2012,
-is now a two-day course delivered both in-person and online
-\cite{lessons-learned,instructor-training,how-to-teach-programming}.
-Not only does it teach good pedagogical practices,
-it serves and an onboarding process to get everyone on the same page
-regarding who lessons are for,
-how they are delivered,
-and how they are maintained.
 
@@
 that can be expanded on or acknowledged to refine the target audience.
+
+\rulemajor{Reduce, re-use, recycle}
+
+Don't create a new lesson if there is an existing one that you can use or contribute to.
+Just as a scholar would not write a paper without a literature review,
+the same holds for lessons.
+Before writing that introduction to the Bash command line,
+for example,
+do a search:
+has anybody else written it?
+Is it complementary to your goals?
+Could it be tweaked or modified to meet your own goals?
+Could your planned lesson be tweaked to compliment the existing lessons so that topics aren't duplicated?
+
+Before re-using content,
+make sure to check that the lesson is licensed in an open manner.
+Both Programming Historian and the Carpentry projects
+use the Creative Commons - Attribution license
+(\url{https://creativecommons.org/licenses/by/4.0/}),
+which allows people to share and adapt material for any purpose
+as long as they cite the original source.
+
+The same questions of re-use come when thinking about recycling content of a lesson,
+such as images, data, figure, or code.
+Does the license cover that as well?
+If not,
+then ask permission,
+just as you would for any other material.
+
+The converse of this rule is to license your own lessons in a similar open manner,
+and to make them discoverable.
+For example,
+when lessons are published (Rule~6),
+make sure that have the usual keywords in their bibliographic entries
+and HTML page headers.
 

### 2017-06-10 paper.tex
Minor edits

@@
 Programming Historian was founded in 2008,
-and has evolved into a collaboratively-edited site providing lessons to humanists. These organizations have been successful in adapting open source development methods
+and has evolved into a collaboratively-edited site providing lessons to humanists.
+These organizations have been successful in adapting open source development methods
 to lesson development and maintenance.
@@
 
-By definition,
-many open education projects satisfy the first criterion,
-but very few satisfy the second:
-while their lessons might occasionally be updated by a small author group,
+Open education projects satisfy the first criterion by definition,
+but very few projects satisfy the second.
+While their lessons might occasionally be updated by a small author group,
 as happens when a new edition of a book is edited and published,
@@
 as maintainers, editors, and reviewers of lessons used by tens of thousands of people.
+
+%%%%% Figure  dissociation 
+\begin{figure}[ht]  % h = here, t = top, b = bottom, p = float
+\includegraphics[width=\linewidth]{figure1}
+\caption{Graphical abstract of 10 simple rules for collaborative lesson development}
+\label{figure1}
+\end{figure}
+%
 
@@
 
-Another example is Software Carpentry's instructor training program.
+An example of how to teach these practices is Software Carpentry's instructor training program.
 First offered in 2012,
@@
 
-It takes a village to ensure
-the steady improvement and adaptation that will give a lesson a long life.
-Ensuring this, however, is not straightforward.
-
-Making the process explicit is key.
+Making the process for contributing to a lesson explicit is
+the key to actually getting contributions.
 New contributors require a straightforward and transparent introduction
@@
 (in open source software development this is known as ``pre-merge review'').
-Yet some of these tools come with a considerable up-front learning curve.
+Providing a gentle on-ramp for new contributors is essential,
+but some tools come with a considerable up-front learning curve.
 GitHub with pull requests is ideal for pre-merge review, for instance,
-but it requires contributors to understand the basics of the git version control software.
+but it requires contributors to know how to use Git,
+which is a famously user-hostile tool.
 Allowing people to edit a Google Doc or Wiki-based editing does not offer pre-merge review,
 but the low barrier to entry can help get conversations started.
-In a pinch,
-work with contributors to find a comfort zone:
-don't let a collaboration flounder on technological choice.
-But make sure not to overwhelm either:
-threaded discussion on paticular topics (e.g. GitHub issues) can help increase the signal-to-noise ratio
+
+The best way to choose tools for managing lessons
+is to look for those that provide a gentle on-ramp for new contributors.
+Ask potential contributors what they are comfortable with
+rather than requiring \emph{them} to come to \emph{you}.
+Remember also that contributing to a lesson is probably not their top priority,
+and look for ways to reduce their cognitive load.
+For example,
+threaded discussion on particular topics (e.g., using GitHub issues)
+can help increase the signal-to-noise ratio
 by reducing long reply-all email excahnges.
 
-Ultimately,
-community dynamics are more important than platform.
-We have seen the 90/9/1 rule:
-90\% of people watch, 9\% discuss, 1\% do.
-This requires a gentle on-ramp for new contributors,
-that we make it easy for people to submit errata and suggestions,
-and that editors may need to do triage to ensure that voices are heard.
-Finally, with so many voices and contributions come caution.
-Working in the open can be great,
+Finally,
+working in the open can be great,
 but can also unintentionally suppress voices.
@@
 If lessons are being released regularly,
-it is worth automating this release process
-with a simple shell script.
+automate the process
+with a shell script or something similar.
 Old versions of lessons should be archived in a discoverable location for reference.

### 2017-06-15 paper.tex
language and citation for git, resolving #7

@@
 but some tools come with a considerable up-front learning curve.
-GitHub with pull requests is ideal for pre-merge review, for instance,
+Housing lessons on GitHub with changes suggested via pull requests is 
+ideal for pre-merge review, for instance,
 but it requires contributors to know how to use Git,
-which is a famously user-hostile tool.
+which has a notoriously steep learning curve \cite{git-survey}.
 Allowing people to edit a Google Doc or Wiki-based editing does not offer pre-merge review,
@@
 
+\bibitem{git-survey}
+GitLab
+\newblock Global Developer Survey.
+\newblock http://get.gitlab.com/global-developer-survey/; 2016.
+
 \end{thebibliography}

### 2017-06-15 paper.tex
adding other systems for lesson development, #8

@@
 can help increase the signal-to-noise ratio
-by reducing long reply-all email excahnges.
+by reducing long reply-all email exchanges. 
+A number of open frameworks are available to facilitate  
+development of new lessons, such as R-based
+learnr (\url{https://rstudio.github.io/learnr}) or 
+GitHub-based Morea (\url{https://morea-framework.github.io}) and 
+DataCamp (\url{https://www.datacamp.com/teach/documentation}).
 

### 2017-06-15 paper.tex
removing extra " from #4

--- 2017-06-15 paper.tex
@@
 For example,
-Programming Historian project labels lessons ``beginner''",
+Programming Historian project labels lessons ``beginner'',
 ``intermediate'',

### 2017-06-15 paper.tex
adding Kate middle initial

@@
 {Rayna~M.~Harris}\textsuperscript{3{\ddag}},
-{Kate~Hertweck}\textsuperscript{4{\ddag}},
+{Kate~L.~Hertweck}\textsuperscript{4{\ddag}},
 {Damien~Irving}\textsuperscript{5{\ddag}},

### 2017-06-17 paper.tex
Fix grammatic usage of open source

@@
 and even more effort to maintain.
-The collaborative code development methods pioneered by the open source community
+The collaborative code development methods pioneered by the open-source community
 offer a way forward,
@@
 By leveraging a community approach, educational resources can be more sustainable, robust, and responsive.
-When lessons are created in an open source community, all lessons are open and accessible,
+When lessons are created in an open-source community, all lessons are open and accessible,
 and they can be continually updated and improved by a community of contibutors.
 
-Despite the proven success of open-source and collaboratively developed lesson plans,
+Despite the proven success of open source and collaboratively developed lesson plans,
 it is surprisingly uncommon in the traditional accademic setting.
@@
 and has evolved into a collaboratively-edited site providing lessons to humanists.
-These organizations have been successful in adapting open source development methods
+These organizations have been successful in adapting open-source development methods
 to lesson development and maintenance.
@@
 prior to their incorporation/merger into the lessons
-(in open source software development this is known as ``pre-merge review'').
+(in open-source software development this is known as ``pre-merge review'').
 Providing a gentle on-ramp for new contributors is essential,
@@
 But sometimes choices must be made.
-The open source software community has wrestled with these issues for three decades,
+The open-source software community has wrestled with these issues for three decades,
 and has evolved some best practices to handle them

### 2017-06-18 paper.tex
Updating affiliation to follow Lyon university guidelines
Tuning: a comma and a missing word

--- 2017-06-15 paper.tex
@@
 \textbf{1} Douglas Mental Health University Institute, McGill University / gdevenyi@gmail.com \\
-\textbf{2} Universit\'{e} Jean Monnet / remi.emonet@univ-st-etienne.fr \\
+\textbf{2} Univ Lyon, UJM-Saint-Etienne, F-42023, France / remi.emonet@univ-st-etienne.fr \\
 \textbf{3} The University of Texas at Austin / rayna.harris@utexas.edu \\
@@
 Lessons don't maintain themselves.
-For technical lessons software versions and dependencies are constantly changing,
+For technical lessons, software versions and dependencies are constantly changing,
 while for more traditional academic lessons the literature is advancing at an ever increasing pace.
@@
 when lessons are published (Rule~6),
-make sure that have the usual keywords in their bibliographic entries
+make sure that they have the usual keywords in their bibliographic entries
 and HTML page headers.

### 2017-06-18 paper.tex
Remove salt and pepper (excess) words

@@
 Despite the proven success of open-source and collaboratively developed lesson plans,
-it is surprisingly uncommon in the traditional accademic setting.
+it is uncommon in the traditional accademic setting.
 Each year, thousands of university lecturers teach subjects ranging from first year biology,
 to graduate-level courses in Indian film.
-Some might use a common textbook written by a single author or two,
+Some use a common textbook written by a single author or two,
 but besides that lecturers develop and improve their course materials in isolation.
@@
 how much valuable time has been wasted,
-and ultimately how the progress of tertiary education has been held back by
+and ultimately how the progress of post-secondary education has been held back by
 the lack of community collaboration and sharing in this process.
 This problem obviously extends beyond the university sector,
-but it is curious that it is endemic in tertiary education,
+but it is curious that it is endemic in post-secondary education,
 given that research depends so critically on collaboration and sharing,
@@
 Programming Historian was founded in 2008,
-and has evolved into a collaboratively-edited site providing lessons to humanists.
+and has evolved into a collaboratively-edited site providing lessons to humanities scholars.
 These organizations have been successful in adapting open source development methods
@@
 Open education projects satisfy the first criterion by definition,
-but very few projects satisfy the second.
-While their lessons might occasionally be updated by a small author group,
+but few projects satisfy the second.
+While their lessons are occasionally updated by a small author group,
 as happens when a new edition of a book is edited and published,
@@
 In particular,
-if lessons reference specific points in the model curriculum guidelines promulgated by many professional societies,
+if lessons reference specific points in the model curriculum guidelines promulgated by professional societies,
 it can be much easier for people to find them and integrate them.
@@
 
-Decades of pedagogical research has yielded many insights into
+Decades of pedagogical research has yielded insights into
 how best to build and deliver lessons \cite{hlw}.
@@
 Not only does it teach good pedagogical practices,
-it serves and an onboarding process to get everyone on the same page
+it serves and an introductory process to get everyone on the same page
 regarding who lessons are for,
@@
 especially if they allow proposed changes to be viewed and discussed
-prior to their incorporation/merger into the lessons
+prior to their incorporation into the lessons
 (in open source software development this is known as ``pre-merge review'').
@@
 but some tools come with a considerable up-front learning curve.
-Housing lessons on GitHub with changes suggested via pull requests is 
+Housing lessons on GitHub (a web-based version control system)
+with changes suggested via pull requests (requests to incorporate changes) is 
 ideal for pre-merge review, for instance,
@@
 If lessons are being released regularly,
-automate the process
-with a shell script or something similar.
+automate the process.
 Old versions of lessons should be archived in a discoverable location for reference.
@@
 to fully understand the frame of reference of their audience.
-For instance, a lesson might inadvertently assume prior knowledge that many learners don't have,
+For instance, a lesson might inadvertently assume prior knowledge that learners don't have,
 which is information that can be collected in a survey.
@@
 Two people with no prior knowledge of a specific subject
-may still be able to move at very different speeds
+may still be able to move at different speeds
 because of different levels of general background knowledge.
@@
 
-If there are several complementary ways to explain something,
-or several points of views that can cohabit respectfully,
+If there are complementary ways to explain something,
+or points of view that can cohabit respectfully,
 it may be possible to present them side by side.
@@
 
-Forking rarely happens in practice.
-When it does,
+These splits rarely happen in practice.
+When they do,
 it is important to remember that we all share the same vision of better lessons, built together.

### 2017-06-18 paper.tex
Small latex formatting fixes

--- 2017-06-18 paper.tex
@@
 \textbf{1} Douglas Mental Health University Institute, McGill University / gdevenyi@gmail.com \\
-\textbf{2} Univ Lyon, UJM-Saint-Etienne, F-42023, France / remi.emonet@univ-st-etienne.fr \\
+\textbf{2} Universit\'{e} Jean Monnet / remi.emonet@univ-st-etienne.fr \\
 \textbf{3} The University of Texas at Austin / rayna.harris@utexas.edu \\
@@
 For example,
-Programming Historian project labels lessons ``beginner'',
-``intermediate'',
+Programming Historian project labels lessons ``beginner,''
+``intermediate,''
 and ``advanced''
@@
 Lessons don't maintain themselves.
-For technical lessons, software versions and dependencies are constantly changing,
+For technical lessons software versions and dependencies are constantly changing,
 while for more traditional academic lessons the literature is advancing at an ever increasing pace.
@@
 such as the full year and month of the release,
-e.g. ``2017.05''.
+e.g. ``2017.05.''
 This is the scheme Software Carpentry has used in
@@
 Both Programming Historian and the Carpentry projects
-use the Creative Commons - Attribution license
+use the Creative Commons -- Attribution license
 (\url{https://creativecommons.org/licenses/by/4.0/}),
@@
 when lessons are published (Rule~6),
-make sure that they have the usual keywords in their bibliographic entries
+make sure that have the usual keywords in their bibliographic entries
 and HTML page headers.

### 2017-06-18 paper.tex
Second round of excess word removal and grammar cleaning

--- 2017-06-18 paper.tex
@@
 equipment or software required,
-how much time learners will actually have:
+how much time learners will have:
 if different contributors believe different things about these,
@@
 
-Lessons don't maintain themselves.
+Lessons do not maintain themselves.
 For technical lessons, software versions and dependencies are constantly changing,
@@
 and contribution, many eyes can keep the lesson usable.
-As we note below,
+As we describe in Rule~5,
 this only works if you make space for
@@
 Less obviously,
-if people have a shared model of how lessons are going to be \emph{used},
+if people have a shared model of how lessons will be \emph{used},
 they are more likely to try to build the same kind of material.
@@
 Making the process for contributing to a lesson explicit is
-the key to actually getting contributions.
+the key to receiving contributions.
 New contributors require a straightforward and transparent introduction
@@
 by reducing long reply-all email exchanges.
-A number of open frameworks are available to facilitate  
+Several open frameworks are available to facilitate
 development of new lessons, such as R-based
@@
 Learners might provide feedback on things like typographical errors,
-clarity/ease of quiz questions and/or the order in which topics are presented,
+clarity/ease of quiz questions or the order in which topics are presented,
 which the instructor can enter into a work-tracking system (e.g., GitHub issues) at the end of class.
@@
 to fully understand the frame of reference of their audience.
-For instance, a lesson might inadvertently assume prior knowledge that learners don't have,
+For instance, a lesson might inadvertently assume prior knowledge that learners do not have,
 which is information that can be collected in a survey.
 If possible, it's a good idea to conduct the post-class survey 30--60 days after the fact.
-This allows people time to reflect,
+This delay allows people time to reflect,
 meaning they are more likely to give accurate feedback on what they learned
@@
 
-Don't create a new lesson if there is an existing one that you can use or contribute to.
+Do not create a new lesson if there is an existing one that you can use or contribute to.
 Just as a scholar would not write a paper without a literature review,
@@
 
-Doing this is a lot of work,
+Doing this is substantial work,
 and maintaining it is even more so,
@@
 it may be possible to present them side by side.
-There are actually good pedagogical reasons to do this
-even if contributors \emph{don't} disagree:
+There are good pedagogical reasons to do this
+even if contributors \emph{do not} disagree:
 weighing alternatives fosters higher-order thinking,
@@
 The open source software community has wrestled with these issues for three decades,
-and has evolved some best practices to handle them
+and has evolved some best practices to address them
 \cite{producing-oss}.

### 2017-06-20 paper.tex

--- 2017-06-18 paper.tex
@@
 For example,
-Programming Historian project labels lessons ``beginner,''
-``intermediate,''
+Programming Historian project labels lessons ``beginner'',
+``intermediate'',
 and ``advanced''
@@
 such as the full year and month of the release,
-e.g. ``2017.05.''
+e.g. ``2017.05''.
 This is the scheme Software Carpentry has used in
@@
 Both Programming Historian and the Carpentry projects
-use the Creative Commons -- Attribution license
+use the Creative Commons - Attribution license
 (\url{https://creativecommons.org/licenses/by/4.0/}),

### 2017-06-20 paper.tex

@@
 and even more effort to maintain.
-The collaborative code development methods pioneered by the open-source community
+The collaborative code development methods pioneered by the open source community
 offer a way forward,
@@
 By leveraging a community approach, educational resources can be more sustainable, robust, and responsive.
-When lessons are created in an open-source community, all lessons are open and accessible,
+When lessons are created in an open source community, all lessons are open and accessible,
 and they can be continually updated and improved by a community of contibutors.
 
-Despite the proven success of open source and collaboratively developed lesson plans,
+Despite the proven success of open-source and collaboratively developed lesson plans,
 it is surprisingly uncommon in the traditional accademic setting.
@@
 Programming Historian was founded in 2008,
-and has evolved into a collaboratively-edited site providing lessons to humanists.
-These organizations have been successful in adapting open-source development methods
+and has evolved into a collaboratively-edited site providing lessons to humanities scholars.
+These organizations have been successful in adapting open source development methods
 to lesson development and maintenance.
@@
 especially if they allow proposed changes to be viewed and discussed
-prior to their incorporation/merger into the lessons
-(in open-source software development this is known as ``pre-merge review'').
+prior to their incorporation into the lessons
+(in open source software development this is known as ``pre-merge review'').
 Providing a gentle on-ramp for new contributors is essential,
 but some tools come with a considerable up-front learning curve.
-Housing lessons on GitHub with changes suggested via pull requests is 
+Housing lessons on GitHub (a web-based version control system)
+with changes suggested via pull requests (requests to incorporate changes) is 
 ideal for pre-merge review, for instance,
@@
 But sometimes choices must be made.
-The open-source software community has wrestled with these issues for three decades,
+The open source software community has wrestled with these issues for three decades,
 and has evolved some best practices to handle them

### 2017-06-20 paper.tex
Callout box with learner profile

--- 2017-06-20 paper.tex
@@
 \usepackage{array}
+
+% frames around boxes
+\usepackage{mdframed}
 
@@
 and any special needs they might have.
+
+\begin{mdframed}
+
+\noindent
+Box 1: Learner Profile
+
+Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
+Other than using Excel, Word, and the Internet,
+Jorge's most significant previous experience with computers is
+helping his sister build a WordPress site for the family business.
+
+Jorge needs to measure properties of soil from nearby farms
+using a handheld device that sends text files to his computer.
+Right now, Jorge has to open each file in Excel,
+crop the first and last points,
+and calculate an average.
+This workshop will show Jorge how to write a little Python program
+to read the data,
+select the right values from each file,
+and calculate the required statistics.
+
+\end{mdframed}
 

### 2017-06-20 paper.tex

--- 2017-06-20 paper.tex
@@
 to fully understand the frame of reference of their audience.
-For instance, a lesson might inadvertently assume prior knowledge that many learners don't have,
+For instance, a lesson might inadvertently assume prior knowledge that learners do not have,
 which is information that can be collected in a survey.
 If possible, it's a good idea to conduct the post-class survey 30--60 days after the fact.
-This allows people time to reflect,
+This delay allows people time to reflect,
 meaning they are more likely to give accurate feedback on what they learned
@@
 
-If there are several complementary ways to explain something,
-or several points of views that can cohabit respectfully,
+If there are complementary ways to explain something,
+or points of view that can cohabit respectfully,
 it may be possible to present them side by side.
-There are actually good pedagogical reasons to do this
-even if contributors \emph{don't} disagree:
+There are good pedagogical reasons to do this
+even if contributors \emph{do not} disagree:
 weighing alternatives fosters higher-order thinking,

### 2017-06-20 paper.tex
Know what 'done' looks like (helps close #4)

@@
 If lessons are being released regularly,
-automate the process.
-Old versions of lessons should be archived in a discoverable location for reference.
+automate the process,
+and archive old versions of lessons in a discoverable location for reference.
+Also make sure that everyone involved knows what ``done'' look like,
+i.e.,
+which outstanding issues have to be addressed
+and how it has to be formatted
+in order for the next release to go out.
+A simple checklist stored with the lesson materials is good enough to start,
+but as time goes by,
+the community may want to use an issue tracking system of some sort
+so that work items can be assigned to specific people
+and then ticked off as they are completed.
 

### 2017-06-20 paper.tex
How to use feedback
Moving figure to end of paper
Adding reference for faculty not having pedagogy training

--- 2017-06-20 paper.tex
@@
 Unfortunately,
-since most college and university faculty have little or no training in education,
+since most college and university faculty have little or no training in education \cite{browning2012},
 this knowledge and expertise is rarely transferred into classroom practice.
@@
 \newblock Software Carpentry: Lessons Learned.
-\newblock F1000Research. 2016;3(62).
-\newblock doi:{10.12688/f1000research.3-62.v2}.
+\newblock F1000Research, 3(62), 2016,
+\newblock 10.12688/f1000research.3-62.v2.
 
 \bibitem{instructor-training}
-Christina Koch and Greg Wilson (eds.)
+Koch C. and Wilson G. (eds.)
 \newblock Software Carpentry: Instructor Training; 2016.
@@
 \bibitem{shell2015}
-Gabriel A.~Devenyi and Christina Koch (eds.):
+Devenyi G.~A. and Koch C. (eds.):
 Software Carpentry: The Unix Shell; 2015.
@@
 \bibitem{shell2017}
-Gabriel A.~Devenyi and Ashwin Srinath (eds.):
+Devenyi G.~A. and Srinath A. (eds.):
 Software Carpentry: The Unix Shell; 2017.
@@
 Caulfield M. Choral Explanations; 2016.
-\newblock https://hapgood.us/2016/05/13/choral-explanations/.
+\newblock \url{https://hapgood.us/2016/05/13/choral-explanations/}.
 
@@
 \newblock Global Developer Survey.
-\newblock http://get.gitlab.com/global-developer-survey/; 2016.
+\newblock \url{http://get.gitlab.com/global-developer-survey/}, 2016.
+
+\bibitem{browning2012}
+Brownell S.~E. and Tanner K.~D.:
+Barriers to faculty pedagogical change: Lack of training, time, incentives, and...tensions with professional identity?
+\newblock CBE Life Sciences Education,
+\newblock 11(4), 2012,
+\newblock 10.1187/cbe.12-09-0163.
 

### 2017-06-29 paper.tex

--- 2017-06-20 paper.tex
@@
 
-\rulemajor{Build community around lessons}
-
-Lessons do not maintain themselves.
-For technical lessons, software versions and dependencies are constantly changing,
-while for more traditional academic lessons the literature is advancing at an ever increasing pace.
-In either case, what was cutting edge in 2017 may be dated
-and less useful in 2018.
-This needs to be clear to contributors,
-so they understand that this lesson is just
-the starting point.
-Sustainability needs to be front of mind.
-
-The focus accordingly needs to be on creating a community.
-Authors cannot be expected to maintain continual vigilance on a lesson,
-but this is necessary if one expects continual use of a lesson!
-By working online, creating opportunities for collaboration
-and contribution, many eyes can keep the lesson usable.
-As we describe in Rule~5,
-this only works if you make space for
-little contributions just as you do for larger ones.
-
 \rulemajor{Build modular lessons that can be re-purposed}
@@
 
+\rulemajor{Build community around lessons}
+
+Encouraging and enabling individual contributions is necessary but not sufficient.
+For technical lessons,
+software versions and dependencies are constantly changing,
+while for more traditional academic lessons the literature is advancing at an ever increasing pace.
+In either case,
+what was cutting edge in 2017 may be dated and less useful in 2018.
+This needs to be clear to contributors,
+so they understand that this lesson is just
+the starting point.
+Sustainability needs to be front of mind.
+
+The focus accordingly needs to be on creating a community in which contributors support each other
+rather than relying on a small group of stewards.
+Authors cannot be expected to maintain continual vigilance on a lesson,
+but this is necessary if one expects continual use of a lesson.
+By working online, creating opportunities for collaboration
+and contribution, many eyes can keep the lesson usable.
+As we describe in Rule~5,
+this only works if you make space for
+little contributions just as you do for larger ones.
+
 \rulemajor{Publish periodically and recognize contributions}
@@
 If lessons are being released regularly,
-automate the process,
-and archive old versions of lessons in a discoverable location for reference.
-Also make sure that everyone involved knows what ``done'' look like,
-i.e.,
-which outstanding issues have to be addressed
-and how it has to be formatted
-in order for the next release to go out.
-A simple checklist stored with the lesson materials is good enough to start,
-but as time goes by,
-the community may want to use an issue tracking system of some sort
-so that work items can be assigned to specific people
-and then ticked off as they are completed.
+automate the process.
+Old versions of lessons should be archived in a discoverable location for reference.
 

### 2017-06-29 paper.tex

--- 2017-06-20 paper.tex
@@
 an ombudsperson is available for private chats and facilitation.
+
+\rulemajor{Build community around lessons}
+
+Encouraging and enabling individual contributions is necessary but not sufficient.
+For technical lessons,
+software versions and dependencies are constantly changing,
+while for more traditional academic lessons the literature is advancing at an ever increasing pace.
+In either case,
+what was cutting edge in 2017 may be dated and less useful in 2018.
+This needs to be clear to contributors,
+so they understand that this lesson is just
+the starting point.
+Sustainability needs to be front of mind.
+
+The focus accordingly needs to be on creating a community in which contributors support each other
+rather than relying on a small group of stewards.
+Authors cannot be expected to maintain continual vigilance on a lesson,
+but this is necessary if one expects continual use of a lesson.
+By working online, creating opportunities for collaboration
+and contribution, many eyes can keep the lesson usable.
+As we describe in Rule~5,
+this only works if you make space for
+little contributions just as you do for larger ones.
 

### 2017-06-29 paper.tex
Fixing wording

@@
 Unfortunately,
-since most college and university faculty have little or no training in education \cite{browning2012},
+since a large number of college and university faculty have little or no training in education \cite{browning2012},
 this knowledge and expertise is rarely transferred into classroom practice.

### 2017-06-29 paper.tex
Merging

@@
 \newcolumntype{+}{!{\vrule width 2pt}}
+
+% frames around boxes
+\usepackage{mdframed}
 
@@
 
+\begin{mdframed}
+
+\noindent
+Box 1: Learner Profile
+
+Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
+While fluent in both Spanish and English,
+he has a hearing disability that sometimes makes it hard for him to understand lectures,
+particularly in noisy environments.
+Other than using Excel, Word, and the Internet,
+Jorge's most significant previous experience with computers is
+helping his sister build a WordPress site for the family business.
+
+Jorge needs to measure properties of soil from nearby farms
+using a handheld device that sends text files to his computer.
+Right now, Jorge has to open each file in Excel,
+crop the first and last points,
+and calculate an average.
+This workshop will show Jorge how to write a little Python program
+to read the data,
+select the right values from each file,
+and calculate the required statistics.
+
+\end{mdframed}
+
 \rulemajor{Build modular lessons that can be re-purposed}
@@
 that can be expanded on or acknowledged to refine the target audience.
+This is one example of a larger guideline:
+when processing feedback,
+separate items that apply to specific lessons (and can guide its future development)
+from those that are better tackled elsewhere
+(e.g., by developing new lessons or through organizational change).
 

### 2017-06-30 paper.tex
Partial rewrite

@@
 
-Lessons take significant effort to build,
-and even more effort to maintain.
-The collaborative code development methods pioneered by the open source community
-offer a way forward,
-allowing us to create lessons which are open,
-accessible,
-and continually updated and improved by a community of contributors.
-This paper provides ten simple rules that can help others create sustainable lessons.
+The collaborative development methods pioneered by the open source software community
+offer a way to create lessons that are open, accessible, and sustainable.
+This paper presents ten simple rules for doing this
+drawn from our experience with several successful projects.
 
@@
 
-The model of collaborative code development presents an
-alternative model to traditional scientific lesson
-development. By leveraging a community approach, educational
-resources can be more sustainable, robust, and responsive.
-These ten simple rules outline best practices for this model of
-collaborative resource development.
+Lessons take significant effort to build and  maintain.
+We have found that
+the collaborative development methods pioneered by the open source software community
+are an effective, economical way to create and sustain lessons
+that can be used by large numbers of people in a wide variety of contexts.
+The ten simple rules outlined in this paper
+summarize the best practices that we have helped evolve
+in several successful open education projects aimed at researchers and research librarians
+in a wide range of disciplines.
 
@@
 
-Lessons take significant time and energy to build and even more effort to maintain.
-Collaborative lesson development refers to the combined efforts of teachers to build and maintain lessons.
-By leveraging a community approach, educational resources can be more sustainable, robust, and responsive.
-When lessons are created in an open source community, all lessons are open and accessible,
-and they can be continually updated and improved by a community of contibutors.
+Lessons take significant effort to build and even more to maintain.
+Most academics do this work on their own,
+but leveraging a community approach
+can make educational resource development more sustainable, robust, and responsive.
+Treating lessons as a community resource
+to be updated, adapted, and improved incrementally,
+can free up valuable time while increasing quality.
 
-Despite the proven success of open-source and collaboratively developed lesson plans,
-it is uncommon in the traditional accademic setting.
-Each year, thousands of university lecturers teach subjects ranging from first year biology,
+Despite the success of openness in software development and the curation of Wikipedia,
+it is an uncommon approach in accademic settings.
+Each year,
+thousands of university lecturers teach subjects ranging from first year biology,
 to graduate-level courses in Indian film.
-Some use a common textbook written by a single author or two,
-but besides that lecturers develop and improve their course materials in isolation.
-It is staggering to think how many wheels are being re-invented,
-how much valuable time has been wasted,
-and ultimately how the progress of post-secondary education has been held back by
-the lack of community collaboration and sharing in this process.
-This problem obviously extends beyond the university sector,
-but it is curious that it is endemic in post-secondary education,
-given that research depends so critically on collaboration and sharing,
-and that most researchers complain about how much time teaching takes away from research.
+Some use textbooks written by one or a few authors,
+but beyond that,
+they develop and maintain their course materials in isolation.
+This is curious given that research depends critically on sharing,
+and that most researchers complain about how much time teaching takes away from research,
+but the sociology and pscyhology behind this blind spot are out of the scope of this paper.
 
-The authors collectively have experience with community-developed lesson plans
-in the context of research computing in the sciences and humanities.
-They primarily draw on their experiences with pedagogical organizations such as
-Software Carpentry and Programming Historian.
+The authors have many years of experience with community-developed lessons
+in the context of research computing in the sciences and humanities
+through organizations like Software Carpentry and Programming Historian.
 Software Carpentry was founded in 1998 to teach scientists basic computing skills,
@@
 and has evolved into a collaboratively-edited site providing lessons to humanities scholars.
-These organizations have been successful in adapting open source development methods
-to lesson development and maintenance.
 Their guiding principles are that lessons should be:
@@
 
-Open education projects satisfy the first criterion by definition,
-but few projects satisfy the second.
-While their lessons are occasionally updated by a small author group,
-as happens when a new edition of a book is edited and published,
+All open education projects satisfy the first criterion by definition,
+but few satisfy the second.
+While their lessons are occasionally updated by a small team
+(as happens when a new edition of a book is edited and published),
 this is not the same as continuous improvement by a large community of contributors.
-The ten simple rules that follow summarize what we have learned about doing that
-as maintainers, editors, and reviewers of lessons used by tens of thousands of people.
-
-%%%%% Figure  dissociation
-\begin{figure}[ht]  % h = here, t = top, b = bottom, p = float
-\includegraphics[width=\linewidth]{figure1}
-\caption{Graphical abstract of 10 simple rules for collaborative lesson development}
-\label{figure1}
-\end{figure}
-%
+The ten simple rules that follow summarize what we have learned about doing that.
 
@@
 or someone in between?
-Prerequisite knowledge,
+If different contributors believe different things about prerequisite knowledge,
 equipment or software required,
-how much time learners will have:
-if different contributors believe different things about these,
-they will find it difficult or impossible to work together.
-
-Thinking systematically about difficulty levels can help manage expectations.
-For example,
-Programming Historian project labels lessons ``beginner'',
-``intermediate'',
-and ``advanced''
-to help authors write at appropriate levels.
-This method works, however, only if there is prior agreement on what those terms mean.
+or how much time learners will have,
+they will find it difficult to work together.
 
@@
 and any special needs they might have.
+This technique is borrowed from user interface design,
+and a typical learner profile is presented in Box~1.
 
-\begin{mdframed}
-
-\noindent
-Box 1: Learner Profile
-
-Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
-While fluent in both Spanish and English,
-he has a hearing disability that sometimes makes it hard for him to understand lectures,
-particularly in noisy environments.
-Other than using Excel, Word, and the Internet,
-Jorge's most significant previous experience with computers is
-helping his sister build a WordPress site for the family business.
-
-Jorge needs to measure properties of soil from nearby farms
-using a handheld device that sends text files to his computer.
-Right now, Jorge has to open each file in Excel,
-crop the first and last points,
-and calculate an average.
-This workshop will show Jorge how to write a little Python program
-to read the data,
-select the right values from each file,
-and calculate the required statistics.
-
-\end{mdframed}
-
-\rulemajor{Build modular lessons that can be re-purposed}
+\rulemajor{Make lessons modular}
 
@@
 so build small chunks that can be re-purposed in many ways.
-A university lecturer in meteorology, for instance,
-might construct a course by bringing together lessons on differential equations,
-fluid mechanics and absorption spectroscopy.
-This task is greatly simplified if existing courses on mathematics,
-physics and chemistry consist of numerous small, discrete lessons,
-as opposed to a few large, monolithic lessons.
+A university lecturer in meteorology,
+for instance,
+might construct a course for their students by bringing together lessons on differential equations,
+fluid mechanics,
+and absorption spectroscopy.
+Creating courses this way shifts the instructor's burden from writing to finding and synthesizing,
+which are easier if lessons clearly define what they cover (Rule~1),
+and if lessons have been designed by people with a shared world-view (Rule~3).
 
-Doing this shifts the instructor's burden from writing to finding and synthesizing,
-both of which are easier if lessons have been designed by people with a shared world-view (Rule~4),
-and if lessons clearly signal what they cover (Rule~1).
-In particular,
-if lessons reference specific points in the model curriculum guidelines promulgated by many professional societies,
-it can be much easier for people to find them and integrate them.
-
-Note also that individual lesson topics hinge on the learners having the proper prerequisite background,
-so this rule further emphasizes the importance of Rule~1.
+One way to achieve this is to take existing courses and break them down into smaller, single-purpose modules
+(a change which has pedagogical and administrative advantages in its own right).
+When this is done,
+these modules can be made more discoverable
+by referencing specific points in the model curricula promulgated by many professional societies.
+Smaller modules are also more approachable for new contributors (Rule~4).
 
@@
 Unfortunately,
-since a large number of college and university faculty have little or no training in education \cite{browning2012},
-this knowledge and expertise is rarely transferred into classroom practice.
+many college and university faculty have little or no training in education \cite{browning2012},
+so this knowledge is rarely applied in the classroom.
 
-Experience shows that even a brief introduction to a handful of key practices
-can help collaborative lesson development in at least three ways.
+Our experience is that even a brief introduction to a few key practices
+helps collaborative lesson development.
 If people have a shared understanding of how lessons should be developed,
@@
 if people have a shared model of how lessons will be \emph{used},
-they are more likely to try to build the same kind of material.
+they are more likely to build reusable material.
 Finally,
@@
 
-An example of a particular lesson development practice is \emph{reverse instructional design}
-\cite{wiggins-mctighe}.
+One popular lesson development methodology is presented in \cite{wiggins-mctighe}.
 When this is used,
-lessons are built by
+lessons are built by:
+
+\begin{enumerate}
+
+  \item
     identifying learning objectives,
+
+  \item
     creating \emph{summative assessments} to determine whether those objectives have been met,
+
+  \item
     designing \emph{formative assessments} to gauge learners' progress
     and give them a chance to practice key skills,
+
+  \item
     putting those formative assessments in order,
+
+  \item
     and only then writing lessons to connect each to the next.
+
+\end{enumerate}
+
+\noindent
 This method is effective in its own right,
-but its greatest benefit is that it gives everyone a common framework
-within which to collaborate.
+but its greatest benefit is that it gives everyone a framework for collaboration.
 
@@
 \cite{lessons-learned,instructor-training,how-to-teach-programming}.
-Not only does it teach good pedagogical practices,
-it serves and an introductory process to get everyone on the same page
-regarding who lessons are for,
+This program teaches good pedagogical practices,
+and introduces everyone who takes it to who Software Carpentry's lessons are for,
 how they are delivered,
 and how they are maintained.
+Largely as a result of this training,
+several hundred people per year now contribute to Software Carpentry's lessons.
 
@@
 
-Making the process for contributing to a lesson explicit is
-the key to receiving contributions.
-New contributors require a straightforward and transparent introduction
-to understand the process of tweaking and adapting lessons.
-Licensing, code of conduct, governance, and contribution procedure
-all need to be explicit rather than implicit
+Making the process for contributing to a lesson simple and transparent
+is the key to receiving contributions.
+Licensing, code of conduct, governance, and the review and publication process
+must all be explicit rather than implicit
 to lower the social barriers to contribution.
@@
 especially if they allow proposed changes to be viewed and discussed
-prior to their incorporation into the lessons
-(in open source software development this is known as ``pre-merge review'').
-Providing a gentle on-ramp for new contributors is essential,
-but some tools come with a considerable up-front learning curve.
-Housing lessons on GitHub (a web-based version control system)
-with changes suggested via pull requests (requests to incorporate changes) is
-ideal for pre-merge review, for instance,
-but it requires contributors to know how to use Git,
+prior to their incorporation into the lessons.
+(In software development this is known as ``pre-merge review''.)
+However,
+some tools that are popular in open source software development have considerable up-front learning costs.
+Portals like GitHub,
+for example,
+support everything that open lesson development needs,
+but require contributors to use Git,
 which has a notoriously steep learning curve \cite{git-survey}.
-Allowing people to edit a Google Doc or Wiki-based editing does not offer pre-merge review,
-but the low barrier to entry can help get conversations started.
+Google Docs and wikis do not support pre-merge review,
+but their low barrier to entry make them more welcoming to newcomers.
 
-The best way to choose tools for managing lessons
-is to look for those that provide a gentle on-ramp for new contributors.
-Ask potential contributors what they are comfortable with
+The best way to choose tools for managing lessons is
+to ask potential contributors what they are comfortable with
 rather than requiring \emph{them} to come to \emph{you}.
@@
 For example,
-threaded discussion on particular topics (e.g., using GitHub issues)
-can help increase the signal-to-noise ratio
+threaded discussion forums can improve the signal-to-noise ratio
 by reducing long reply-all email exchanges.
-Several open frameworks are available to facilitate
-development of new lessons, such as R-based
-learnr (\url{https://rstudio.github.io/learnr}) or
-GitHub-based Morea (\url{https://morea-framework.github.io}) and
-DataCamp (\url{https://www.datacamp.com/teach/documentation}).
+Several open frameworks are available to facilitate development of new lessons,
+such as R-based learnr (\url{https://rstudio.github.io/learnr})
+or GitHub-based Morea (\url{https://morea-framework.github.io})
+and DataCamp's templates (\url{https://www.datacamp.com/teach/documentation}).
+
+\rulemajor{Build community around lessons}
+
+Software versions and dependencies are constantly changing,
+while the academic literature is advancing at an ever-increasing pace.
+As a result,
+what is cutting edge one year may be out of date the next and simply wrong the year after.
+Collaborative lesson development groups must therefore focus
+on creating a community in which contributors support each other
+rather than on relying on a small group of stewards.
+Authors cannot be expected to maintain continual vigilance on a lesson,
+but this is necessary for continual use.
+
+A key part of doing this is to create opportunities for legitimate peripheral participation \cite{FIXME}.
+Curating a list of small tasks that newcomers can easily tackle,
+encouraging them to give feedback on proposed changes,
+or asking them to add new exercises and tweak diagrams and references
+can all provide an on-ramp for people who might question their authority or ability to change the main body of a lesson.
+Equally,
+acknowledging all contributions,
+however small,
+gives new contributors an early reward for taking part.
 
@@
 but can also unintentionally suppress voices.
-At Programming Historian,
-for example,
-an ombudsperson is available for private chats and facilitation.
-
-\rulemajor{Build community around lessons}
-
-Encouraging and enabling individual contributions is necessary but not sufficient.
-For technical lessons,
-software versions and dependencies are constantly changing,
-while for more traditional academic lessons the literature is advancing at an ever increasing pace.
-In either case,
-what was cutting edge in 2017 may be dated and less useful in 2018.
-This needs to be clear to contributors,
-so they understand that this lesson is just
-the starting point.
-Sustainability needs to be front of mind.
-
-The focus accordingly needs to be on creating a community in which contributors support each other
-rather than relying on a small group of stewards.
-Authors cannot be expected to maintain continual vigilance on a lesson,
-but this is necessary if one expects continual use of a lesson.
-By working online, creating opportunities for collaboration
-and contribution, many eyes can keep the lesson usable.
-As we describe in Rule~5,
-this only works if you make space for
-little contributions just as you do for larger ones.
+Programming Historian makes an ombudsperson is available for private chats and facilitation
+to ensure that no one is excluded.
 
@@
 Like software,
-lessons should have releases of fixed content
-so that learners or instructors who may wish to use the material have a stable version to refer to
-for the duration of their use.
-These releases should be periodic
-so that improvements and adjustments are made available
-for new learners and instructors just starting their use.
-Periodic releases are also essential
-for enabling recognition of the contribution of authors and maintainers.
+specific versions of lessons should be published or released periodically
+so that learners or instructors have something stable to refer to for the duration of their use.
+Periodic releases also provide an opportunity
+for recognizing the contributions of new authors and maintainers.
 
@@
 it is important to publish lessons in ways that traditional academic systems can digest.
-One is to give particular releases of lessons
-DOIs supplied by providers such as Zenodo (\url{https://zenodo.org/})
+One is to give releases DOIs supplied by providers such as Zenodo (\url{https://zenodo.org/})
 or DataCite (\url{https://www.datacite.org/}).
@@
 Each time the lesson is published,
-names and identifiers (e.g. ORCIDs (\url{https://orcid.org}))
+names (and identifiers such as ORCIDs (\url{https://orcid.org}))
 should be gathered for all contributors.
@@
 by fixing outstanding issues and merging contributions.
-Version control helps a lot in continuously maintaining a list of contributors
-but also in remembering which version is used for release
+Version control automatically maintains a list of contributors,
+and can also be used to track what content is in what release
 (e.g., using branches or tags).
-Lesson releases should use a consistent naming scheme,
-such as the full year and month of the release,
-e.g. ``2017.05''.
-This is the scheme Software Carpentry has used in
-its releases of its lessons \cite{shell2015,shell2017}.
+Lesson releases should use a consistent naming scheme;
+Software Carpentry has used the year and month of release
+(e.g., ``2017.05'')
+in its releases \cite{shell2015,shell2017}.
 
 If lessons are being released regularly,
-automate the process,
-and archive old versions of lessons in a discoverable location for reference.
+automate the process
+and archive old versions in a discoverable location.
 Also make sure that everyone involved knows what ``done'' look like,
@@
 
-The purpose of feedback is to guide lesson development
-so that authors aren't designing and arguing in a vacuum.
-What people immersed in the lessons think needs fixing
-can often differ from what learners think.
+What people immersed in developing lessons think needs fixing
+can easily differ from what learners think.
+It is therefore critical to gather and act on feedback at several scales
+to check assumptions and stay on course.
 
 Micro-scale feedback can be gathered by an instructor while teaching a particular lesson.
-Learners might provide feedback on things like typographical errors,
-clarity/ease of quiz questions or the order in which topics are presented,
-which the instructor can enter into a work-tracking system (e.g., GitHub issues) at the end of class.
+Learners can provide feedback on everything from typographical errors
+and the clarity of quiz questions
+to the order in which topics are presented,
+all of which the instructor should record at the end of each class
+in some shared location (such as a Google Doc or GitHub issues).
 As well as encouraging direct verbal feedback,
 it's a good idea to provide learners with a means to provide feedback anonymously during class
-(e.g., on small pieces of paper like sticky notes).
+(e.g., on small pieces of paper like sticky notes or through anonymous surveys).
 
 Pre- and post-class surveys can be used to discover larger macro-scale issues.
-These issues often relate to the fact that it can be difficult for lesson developers
-to fully understand the frame of reference of their audience.
-For instance, a lesson might inadvertently assume prior knowledge that learners do not have,
-which is information that can be collected in a survey.
-If possible, it's a good idea to conduct the post-class survey 30--60 days after the fact.
+These often relate to the difficulty lesson developers may have
+fully understanding the frame of reference of their audience,
+e.g.,
+assuming prior knowledge that learners do not have.
+If possible,
+post-class surveys should be conducted 30--60 days after class.
 This delay allows people time to reflect,
@@
 
-Pre- and post-class surveys also are essential in focusing in on the audience for a given lesson.
-Referring back to Rule~1,
-it can sometimes be difficult for lesson developers to understand the frame of a given learner,
-so surveys (particularly post-class surveys) can reveal hidden assumed knowledge
-that can be expanded on or acknowledged to refine the target audience.
-This is one example of a larger guideline:
-when processing feedback,
-separate items that apply to specific lessons (and can guide its future development)
-from those that are better tackled elsewhere
-(e.g., by developing new lessons or through organizational change).
-
 \rulemajor{Reduce, re-use, recycle}
 
-Do not create a new lesson if there is an existing one that you can use or contribute to.
 Just as a scholar would not write a paper without a literature review,
-the same holds for lessons.
-Before writing that introduction to the Bash command line,
-for example,
-do a search:
-has anybody else written it?
-Is it complementary to your goals?
-Could it be tweaked or modified to meet your own goals?
-Could your planned lesson be tweaked to compliment the existing lessons so that topics aren't duplicated?
+an instructor should not create a new lesson if there is an existing one they could use or contribute to.
+A short online search can discover if someone has written what you need,
+whether it is complementary to your goals,
+and if it can be tweaked or modified to meet your needs.
 
 Before re-using content,
-make sure to check that the lesson is licensed in an open manner.
+make sure to check its license.
 Both Programming Historian and the Carpentry projects
-use the Creative Commons - Attribution license
+use the Creative Commons--Attribution license
 (\url{https://creativecommons.org/licenses/by/4.0/}),
@@
 as long as they cite the original source.
+Other Creative Commons licenses may restrict commercial use
+and/or creation of derivative materials.
 
@@
 
+%%%%% Figure  dissociation
+\begin{figure}[ht]  % h = here, t = top, b = bottom, p = float
+\includegraphics[width=\linewidth]{figure1}
+\caption{Graphical abstract of 10 simple rules for collaborative lesson development}
+\label{figure1}
+\end{figure}
+%
+
+\begin{mdframed}
+
+\noindent
+Box 1: Learner Profile
+
+Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
+While fluent in both Spanish and English,
+he has a hearing disability that sometimes makes it hard for him to understand lectures,
+particularly in noisy environments.
+Other than using Excel, Word, and the Internet,
+Jorge's most significant previous experience with computers is
+helping his sister build a WordPress site for the family business.
+
+Jorge needs to measure properties of soil from nearby farms
+using a handheld device that sends text files to his computer.
+Right now, Jorge has to open each file in Excel,
+crop the first and last points,
+and calculate an average.
+This workshop will show Jorge how to write a little Python program
+to read the data,
+select the right values from each file,
+and calculate the required statistics.
+
+\end{mdframed}
+
 \end{document}

### 2017-06-30 paper.tex
Revising paper

@@
 % \newcommand{\ipsum}{{\bf IPSUM}}
-\newcommand{\rulemajor}[1]{\section{#1}}
+\newcommand{\rulemajor}[2]{\section{#1}\label{#2}}
+\newcommand{\ruleref}[1]{Rule~\ref{#1}}
 
@@
 
-\rulemajor{Clarify your audience}
+\rulemajor{Clarify your audience}{audience}
 
@@
 
-\rulemajor{Make lessons modular}
+\rulemajor{Make lessons modular}{modular}
 
@@
 Creating courses this way shifts the instructor's burden from writing to finding and synthesizing,
-which are easier if lessons clearly define what they cover (Rule~1),
-and if lessons have been designed by people with a shared world-view (Rule~3).
+which are easier if lessons clearly define what they cover (\ruleref{audience}),
+and if lessons have been designed by people with a shared world-view (\ruleref{practices}).
 
@@
 by referencing specific points in the model curricula promulgated by many professional societies.
-Smaller modules are also more approachable for new contributors (Rule~4).
+Smaller modules are also more approachable for new contributors (\ruleref{empower}).
 
-\rulemajor{Teach best practices for lesson development}
+\rulemajor{Teach best practices for lesson development}{practices}
 
@@
 
-\rulemajor{Encourage and empower contributors}
+\rulemajor{Encourage and empower contributors}{empower}
 
@@
 
-\rulemajor{Build community around lessons}
+\rulemajor{Build community around lessons}{community}
 
@@
 
-A key part of doing this is to create opportunities for legitimate peripheral participation \cite{FIXME}.
+A key part of doing this is to create opportunities for legitimate peripheral participation.
 Curating a list of small tasks that newcomers can easily tackle,
@@
 
-\rulemajor{Publish periodically and recognize contributions}
+\rulemajor{Publish periodically and recognize contributions}{publish}
 
@@
 
-\rulemajor{Evaluate lessons at several scales}
+\rulemajor{Evaluate lessons at several scales}{evaluate}
 
@@
 
-Pre- and post-class surveys can be used to discover larger macro-scale issues.
-These often relate to the difficulty lesson developers may have
-fully understanding the frame of reference of their audience,
+Pre- and post-class surveys should be used to uncover larger issues,
+particularly those arising from developers not fully understanding their audience,
 e.g.,
 assuming prior knowledge that learners do not have.
-If possible,
-post-class surveys should be conducted 30--60 days after class.
-This delay allows people time to reflect,
-meaning they are more likely to give accurate feedback on what they learned
+Post-class surveys are most effective when conducted 30--90 days after class;
+this gives people time to reflect,
+so their feedback will more accurately reflect what they learned
 rather than how entertained they were.
 
-\rulemajor{Reduce, re-use, recycle}
+\rulemajor{Reduce, re-use, recycle}{rrr}
 
@@
 
-The same questions of re-use come when thinking about recycling content of a lesson,
+The question of licensing also arises when recycling lesson components
 such as images, data, figure, or code.
-Does the license cover that as well?
-If not,
-then ask permission,
-just as you would for any other material.
+If the license does not cover them explicitly,
+ask permission as you would for any other academic material.
 
-The converse of this rule is to license your own lessons in a similar open manner,
-and to make them discoverable.
+The converse of this rule is to make the license for your lessons explicit and discoverable.
 For example,
-when lessons are published (Rule~6),
-make sure that they have the usual keywords in their bibliographic entries
+when lessons are published (\ruleref{publish}),
+make sure that that keywords such as ``CC-BY'' appear in their bibliography entries
 and HTML page headers.
 
-\rulemajor{Link to other resources}
+\rulemajor{Link to other resources}{link}
 
-Most learners are unlikely to absorb everything there is to know about a topic
-from your lesson alone.
+Learners are unlikely to absorb everything they need to know about a topic from your lesson alone.
 This is partly a matter of scope---any interesting subject is too large
@@
 As Caulfield has argued \cite{choral-explanations},
-the best resources provide a chorus of explanations
+the best way to use the Internet is to provide a chorus of explanations
 that offer many angles and approaches for any given topic,
-each of which may be the best fit for a different subset of people.
+each of which may be the best fit for a different set of needs.
 
-Find these resources and direct learners to them at strategic points.
-These resources may include textbooks,
+Collaboratively-developed lessons should direct learners to these resources at strategic points.
+Textbooks,
 technical documentation,
 videos,
-and web pages;
+web pages,
+threads on Quora or mailing lists:
 if a community or discussion forum exists for the topic,
-be sure to include them as well.
+it is worth including.
 
@@
 and maintaining it is even more so,
-which makes building community around lessons (Rule~2) all the more important.
+which makes building community around lessons (\ruleref{community}) all the more important.
 In particular,
@@
 they should both be able to propose updates, corrections, and additions to lessons,
-and know that they are encouraged to do so (Rule~5).
+and know that they are encouraged to do so (\ruleref{empower}).
 
-\rulemajor{You can't please everyone}
+\rulemajor{You can't please everyone}{everyone}
 
@@
 Similarly,
-lessons on ecology for learners in Utah and Florida
+lessons on ecology for learners in Utah and Vietnam
 will probably be more relatable if they use different examples.
-Equally,
+A community may therefore maintain several differently-oriented or differently-paced lessons
+on a single topic,
+just as programming languages provide several different libraries for doing the same general thing
+with different levels of performance and complexity.
+
+Similarly,
 no lesson development community can serve all purposes.
@@
 while others may prefer a ``measure twice, cut once'' approach.
-
 If there are complementary ways to explain something,
@@
 even if contributors \emph{do not} disagree:
-weighing alternatives fosters higher-order thinking,
-and as Caulfield has written,
-the most popular online question and answer sites
-are successful in part because they present a chorus of explanations
-geared at different levels and needs
-rather than a single one \cite{choral-explanations}.
+weighing alternatives fosters higher-order thinking.
 
@@
 \cite{producing-oss}.
-As discussed in Rule~5,
+As discussed in \ruleref{empower},
 the first step is to have a clear governance structure and a clear, permissive license.
 Minor disagreements should be discussed openly and respectfully.
-If they cannot be resolved---i.e., if they turn out not to be so minor after all---then
+If they turn out not to be so minor after all,
 contributors should split off and evolve the lesson in the way they see best.
@@
 % step-by-step instructions.
+
+\newpage
 
@@
 
+\newpage
+
 %%%%% Figure  dissociation

### 2017-06-30 paper.tex
Creating home page version of paper

@@
 Unfortunately,
-many college and university faculty have little or no training in education \cite{browning2012},
+many college and university faculty have little or no training in education \cite{brownell},
 so this knowledge is rarely applied in the classroom.
@@
 
-\bibitem{browning2012}
+\bibitem{brownell}
 Brownell S.~E. and Tanner K.~D.:

### 2017-07-01 paper.tex
tidying up

@@
 Several open frameworks are available to facilitate development of new lessons,
-such as R-based learnr (\url{https://rstudio.github.io/learnr})
-or GitHub-based Morea (\url{https://morea-framework.github.io})
+such as learnr (\url{https://rstudio.github.io/learnr}),
+Morea (\url{https://morea-framework.github.io}),
 and DataCamp's templates (\url{https://www.datacamp.com/teach/documentation}).

### 2017-07-03 paper.tex
adding in-text figure refs

--- 2017-06-30 paper.tex
@@
 % \newcommand{\ipsum}{{\bf IPSUM}}
-\newcommand{\rulemajor}[2]{\section{#1}\label{#2}}
-\newcommand{\ruleref}[1]{Rule~\ref{#1}}
+\newcommand{\rulemajor}[1]{\section{#1}}
 
@@
 
-The collaborative development methods pioneered by the open source software community
-offer a way to create lessons that are open, accessible, and sustainable.
-This paper presents ten simple rules for doing this
-drawn from our experience with several successful projects.
+Lessons take significant effort to build,
+and even more effort to maintain.
+The collaborative code development methods pioneered by the open source community
+offer a way forward,
+allowing us to create lessons which are open,
+accessible,
+and continually updated and improved by a community of contributors.
+This paper provides ten simple rules that can help others create sustainable lessons (Fig. 1).
 
@@
 
-Lessons take significant effort to build and  maintain.
-We have found that
-the collaborative development methods pioneered by the open source software community
-are an effective, economical way to create and sustain lessons
-that can be used by large numbers of people in a wide variety of contexts.
-The ten simple rules outlined in this paper
-summarize the best practices that we have helped evolve
-in several successful open education projects aimed at researchers and research librarians
-in a wide range of disciplines.
+The model of collaborative code development presents an
+alternative model to traditional scientific lesson
+development. By leveraging a community approach, educational
+resources can be more sustainable, robust, and responsive.
+These ten simple rules outline best practices for this model of
+collaborative resource development.
 
@@
 
-Lessons take significant effort to build and even more to maintain.
-Most academics do this work on their own,
-but leveraging a community approach
-can make educational resource development more sustainable, robust, and responsive.
-Treating lessons as a community resource
-to be updated, adapted, and improved incrementally,
-can free up valuable time while increasing quality.
+Lessons take significant time and energy to build and even more effort to maintain.
+Collaborative lesson development refers to the combined efforts of teachers to build and maintain lessons.
+By leveraging a community approach, educational resources can be more sustainable, robust, and responsive.
+When lessons are created in an open source community, all lessons are open and accessible,
+and they can be continually updated and improved by a community of contibutors.
 
-Despite the success of openness in software development and the curation of Wikipedia,
-it is an uncommon approach in accademic settings.
-Each year,
-thousands of university lecturers teach subjects ranging from first year biology,
+Despite the proven success of open-source and collaboratively developed lesson plans,
+it is uncommon in the traditional accademic setting.
+Each year, thousands of university lecturers teach subjects ranging from first year biology,
 to graduate-level courses in Indian film.
-Some use textbooks written by one or a few authors,
-but beyond that,
-they develop and maintain their course materials in isolation.
-This is curious given that research depends critically on sharing,
-and that most researchers complain about how much time teaching takes away from research,
-but the sociology and pscyhology behind this blind spot are out of the scope of this paper.
+Some use a common textbook written by a single author or two,
+but besides that lecturers develop and improve their course materials in isolation.
+It is staggering to think how many wheels are being re-invented,
+how much valuable time has been wasted,
+and ultimately how the progress of post-secondary education has been held back by
+the lack of community collaboration and sharing in this process.
+This problem obviously extends beyond the university sector,
+but it is curious that it is endemic in post-secondary education,
+given that research depends so critically on collaboration and sharing,
+and that most researchers complain about how much time teaching takes away from research.
 
-The authors have many years of experience with community-developed lessons
-in the context of research computing in the sciences and humanities
-through organizations like Software Carpentry and Programming Historian.
+The authors collectively have experience with community-developed lesson plans
+in the context of research computing in the sciences and humanities.
+They primarily draw on their experiences with pedagogical organizations such as
+Software Carpentry and Programming Historian.
 Software Carpentry was founded in 1998 to teach scientists basic computing skills,
@@
 and has evolved into a collaboratively-edited site providing lessons to humanities scholars.
+These organizations have been successful in adapting open source development methods
+to lesson development and maintenance.
 Their guiding principles are that lessons should be:
@@
 
-All open education projects satisfy the first criterion by definition,
-but few satisfy the second.
-While their lessons are occasionally updated by a small team
-(as happens when a new edition of a book is edited and published),
+Open education projects satisfy the first criterion by definition,
+but few projects satisfy the second.
+While their lessons are occasionally updated by a small author group,
+as happens when a new edition of a book is edited and published,
 this is not the same as continuous improvement by a large community of contributors.
-The ten simple rules that follow summarize what we have learned about doing that.
+The ten simple rules that follow summarize what we have learned about doing that
+as maintainers, editors, and reviewers of lessons used by tens of thousands of people (Fig. 1, 2).
 
-\rulemajor{Clarify your audience}{audience}
+%%%%% Figure  dissociation
+\begin{figure}[ht]  % h = here, t = top, b = bottom, p = float
+\includegraphics[width=\linewidth]{figure1}
+\caption{Graphical abstract of 10 simple rules for collaborative lesson development}
+\label{figure1}
+\end{figure}
+%
+
+\rulemajor{Clarify your audience}
 
@@
 or someone in between?
-If different contributors believe different things about prerequisite knowledge,
+Prerequisite knowledge,
 equipment or software required,
-or how much time learners will have,
-they will find it difficult to work together.
+how much time learners will have:
+if different contributors believe different things about these,
+they will find it difficult or impossible to work together.
+
+Thinking systematically about difficulty levels can help manage expectations.
+For example,
+Programming Historian project labels lessons ``beginner'',
+``intermediate'',
+and ``advanced''
+to help authors write at appropriate levels.
+This method works, however, only if there is prior agreement on what those terms mean.
 
@@
 and any special needs they might have.
-This technique is borrowed from user interface design,
-and a typical learner profile is presented in Box~1.
 
-\rulemajor{Make lessons modular}{modular}
+\begin{mdframed}
+
+\noindent
+Box 1: Learner Profile
+
+Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
+While fluent in both Spanish and English,
+he has a hearing disability that sometimes makes it hard for him to understand lectures,
+particularly in noisy environments.
+Other than using Excel, Word, and the Internet,
+Jorge's most significant previous experience with computers is
+helping his sister build a WordPress site for the family business.
+
+Jorge needs to measure properties of soil from nearby farms
+using a handheld device that sends text files to his computer.
+Right now, Jorge has to open each file in Excel,
+crop the first and last points,
+and calculate an average.
+This workshop will show Jorge how to write a little Python program
+to read the data,
+select the right values from each file,
+and calculate the required statistics.
+
+\end{mdframed}
+
+\rulemajor{Build modular lessons that can be re-purposed}
 
@@
 so build small chunks that can be re-purposed in many ways.
-A university lecturer in meteorology,
-for instance,
-might construct a course for their students by bringing together lessons on differential equations,
-fluid mechanics,
-and absorption spectroscopy.
-Creating courses this way shifts the instructor's burden from writing to finding and synthesizing,
-which are easier if lessons clearly define what they cover (\ruleref{audience}),
-and if lessons have been designed by people with a shared world-view (\ruleref{practices}).
+A university lecturer in meteorology, for instance,
+might construct a course by bringing together lessons on differential equations,
+fluid mechanics and absorption spectroscopy.
+This task is greatly simplified if existing courses on mathematics,
+physics and chemistry consist of numerous small, discrete lessons,
+as opposed to a few large, monolithic lessons.
 
-One way to achieve this is to take existing courses and break them down into smaller, single-purpose modules
-(a change which has pedagogical and administrative advantages in its own right).
-When this is done,
-these modules can be made more discoverable
-by referencing specific points in the model curricula promulgated by many professional societies.
-Smaller modules are also more approachable for new contributors (\ruleref{empower}).
+Doing this shifts the instructor's burden from writing to finding and synthesizing,
+both of which are easier if lessons have been designed by people with a shared world-view (Rule~4),
+and if lessons clearly signal what they cover (Rule~1).
+In particular,
+if lessons reference specific points in the model curriculum guidelines promulgated by many professional societies,
+it can be much easier for people to find them and integrate them.
 
-\rulemajor{Teach best practices for lesson development}{practices}
+Note also that individual lesson topics hinge on the learners having the proper prerequisite background,
+so this rule further emphasizes the importance of Rule~1.
+
+\rulemajor{Teach best practices for lesson development}
 
@@
 Unfortunately,
-many college and university faculty have little or no training in education \cite{browning2012},
-so this knowledge is rarely applied in the classroom.
+since a large number of college and university faculty have little or no training in education \cite{browning2012},
+this knowledge and expertise is rarely transferred into classroom practice.
 
-Our experience is that even a brief introduction to a few key practices
-helps collaborative lesson development.
+Experience shows that even a brief introduction to a handful of key practices
+can help collaborative lesson development in at least three ways.
 If people have a shared understanding of how lessons should be developed,
@@
 if people have a shared model of how lessons will be \emph{used},
-they are more likely to build reusable material.
+they are more likely to try to build the same kind of material.
 Finally,
@@
 
-One popular lesson development methodology is presented in \cite{wiggins-mctighe}.
+An example of a particular lesson development practice is \emph{reverse instructional design}
+\cite{wiggins-mctighe}.
 When this is used,
-lessons are built by:
-
-\begin{enumerate}
-
-  \item
+lessons are built by
 identifying learning objectives,
-
-  \item
 creating \emph{summative assessments} to determine whether those objectives have been met,
-
-  \item
 designing \emph{formative assessments} to gauge learners' progress
 and give them a chance to practice key skills,
-
-  \item
 putting those formative assessments in order,
-
-  \item
 and only then writing lessons to connect each to the next.
-
-\end{enumerate}
-
-\noindent
 This method is effective in its own right,
-but its greatest benefit is that it gives everyone a framework for collaboration.
+but its greatest benefit is that it gives everyone a common framework
+within which to collaborate.
 
@@
 \cite{lessons-learned,instructor-training,how-to-teach-programming}.
-This program teaches good pedagogical practices,
-and introduces everyone who takes it to who Software Carpentry's lessons are for,
+Not only does it teach good pedagogical practices,
+it serves and an introductory process to get everyone on the same page
+regarding who lessons are for,
 how they are delivered,
 and how they are maintained.
-Largely as a result of this training,
-several hundred people per year now contribute to Software Carpentry's lessons.
 
-\rulemajor{Encourage and empower contributors}{empower}
+\rulemajor{Encourage and empower contributors}
 
-Making the process for contributing to a lesson simple and transparent
-is the key to receiving contributions.
-Licensing, code of conduct, governance, and the review and publication process
-must all be explicit rather than implicit
+Making the process for contributing to a lesson explicit is
+the key to receiving contributions.
+New contributors require a straightforward and transparent introduction
+to understand the process of tweaking and adapting lessons.
+Licensing, code of conduct, governance, and contribution procedure
+all need to be explicit rather than implicit
 to lower the social barriers to contribution.
@@
 especially if they allow proposed changes to be viewed and discussed
-prior to their incorporation into the lessons.
-(In software development this is known as ``pre-merge review''.)
-However,
-some tools that are popular in open source software development have considerable up-front learning costs.
-Portals like GitHub,
-for example,
-support everything that open lesson development needs,
-but require contributors to use Git,
+prior to their incorporation into the lessons
+(in open source software development this is known as ``pre-merge review'').
+Providing a gentle on-ramp for new contributors is essential,
+but some tools come with a considerable up-front learning curve.
+Housing lessons on GitHub (a web-based version control system)
+with changes suggested via pull requests (requests to incorporate changes) is
+ideal for pre-merge review, for instance,
+but it requires contributors to know how to use Git,
 which has a notoriously steep learning curve \cite{git-survey}.
-Google Docs and wikis do not support pre-merge review,
-but their low barrier to entry make them more welcoming to newcomers.
+Allowing people to edit a Google Doc or Wiki-based editing does not offer pre-merge review,
+but the low barrier to entry can help get conversations started.
 
-The best way to choose tools for managing lessons is
-to ask potential contributors what they are comfortable with
+The best way to choose tools for managing lessons
+is to look for those that provide a gentle on-ramp for new contributors.
+Ask potential contributors what they are comfortable with
 rather than requiring \emph{them} to come to \emph{you}.
@@
 For example,
-threaded discussion forums can improve the signal-to-noise ratio
+threaded discussion on particular topics (e.g., using GitHub issues)
+can help increase the signal-to-noise ratio
 by reducing long reply-all email exchanges.
-Several open frameworks are available to facilitate development of new lessons,
-such as R-based learnr (\url{https://rstudio.github.io/learnr})
-or GitHub-based Morea (\url{https://morea-framework.github.io})
-and DataCamp's templates (\url{https://www.datacamp.com/teach/documentation}).
-
-\rulemajor{Build community around lessons}{community}
-
-Software versions and dependencies are constantly changing,
-while the academic literature is advancing at an ever-increasing pace.
-As a result,
-what is cutting edge one year may be out of date the next and simply wrong the year after.
-Collaborative lesson development groups must therefore focus
-on creating a community in which contributors support each other
-rather than on relying on a small group of stewards.
-Authors cannot be expected to maintain continual vigilance on a lesson,
-but this is necessary for continual use.
-
-A key part of doing this is to create opportunities for legitimate peripheral participation.
-Curating a list of small tasks that newcomers can easily tackle,
-encouraging them to give feedback on proposed changes,
-or asking them to add new exercises and tweak diagrams and references
-can all provide an on-ramp for people who might question their authority or ability to change the main body of a lesson.
-Equally,
-acknowledging all contributions,
-however small,
-gives new contributors an early reward for taking part.
+Several open frameworks are available to facilitate
+development of new lessons, such as R-based
+learnr (\url{https://rstudio.github.io/learnr}) or
+GitHub-based Morea (\url{https://morea-framework.github.io}) and
+DataCamp (\url{https://www.datacamp.com/teach/documentation}).
 
@@
 but can also unintentionally suppress voices.
-Programming Historian makes an ombudsperson is available for private chats and facilitation
-to ensure that no one is excluded.
+At Programming Historian,
+for example,
+an ombudsperson is available for private chats and facilitation.
 
-\rulemajor{Publish periodically and recognize contributions}{publish}
+\rulemajor{Build community around lessons}
+
+Encouraging and enabling individual contributions is necessary but not sufficient.
+For technical lessons,
+software versions and dependencies are constantly changing,
+while for more traditional academic lessons the literature is advancing at an ever increasing pace.
+In either case,
+what was cutting edge in 2017 may be dated and less useful in 2018.
+This needs to be clear to contributors,
+so they understand that this lesson is just
+the starting point.
+Sustainability needs to be front of mind.
+
+The focus accordingly needs to be on creating a community in which contributors support each other
+rather than relying on a small group of stewards.
+Authors cannot be expected to maintain continual vigilance on a lesson,
+but this is necessary if one expects continual use of a lesson.
+By working online, creating opportunities for collaboration
+and contribution, many eyes can keep the lesson usable.
+As we describe in Rule~5,
+this only works if you make space for
+little contributions just as you do for larger ones.
+
+\rulemajor{Publish periodically and recognize contributions}
 
 Like software,
-specific versions of lessons should be published or released periodically
-so that learners or instructors have something stable to refer to for the duration of their use.
-Periodic releases also provide an opportunity
-for recognizing the contributions of new authors and maintainers.
+lessons should have releases of fixed content
+so that learners or instructors who may wish to use the material have a stable version to refer to
+for the duration of their use.
+These releases should be periodic
+so that improvements and adjustments are made available
+for new learners and instructors just starting their use.
+Periodic releases are also essential
+for enabling recognition of the contribution of authors and maintainers.
 
@@
 it is important to publish lessons in ways that traditional academic systems can digest.
-One is to give releases DOIs supplied by providers such as Zenodo (\url{https://zenodo.org/})
+One is to give particular releases of lessons
+DOIs supplied by providers such as Zenodo (\url{https://zenodo.org/})
 or DataCite (\url{https://www.datacite.org/}).
@@
 Each time the lesson is published,
-names (and identifiers such as ORCIDs (\url{https://orcid.org}))
+names and identifiers (e.g. ORCIDs (\url{https://orcid.org}))
 should be gathered for all contributors.
@@
 by fixing outstanding issues and merging contributions.
-Version control automatically maintains a list of contributors,
-and can also be used to track what content is in what release
+Version control helps a lot in continuously maintaining a list of contributors
+but also in remembering which version is used for release
 (e.g., using branches or tags).
-Lesson releases should use a consistent naming scheme;
-Software Carpentry has used the year and month of release
-(e.g., ``2017.05'')
-in its releases \cite{shell2015,shell2017}.
+Lesson releases should use a consistent naming scheme,
+such as the full year and month of the release,
+e.g. ``2017.05''.
+This is the scheme Software Carpentry has used in
+its releases of its lessons \cite{shell2015,shell2017}.
 
 If lessons are being released regularly,
-automate the process
-and archive old versions in a discoverable location.
+automate the process,
+and archive old versions of lessons in a discoverable location for reference.
 Also make sure that everyone involved knows what ``done'' look like,
@@
 
-\rulemajor{Evaluate lessons at several scales}{evaluate}
+\rulemajor{Evaluate lessons at several scales}
 
-What people immersed in developing lessons think needs fixing
-can easily differ from what learners think.
-It is therefore critical to gather and act on feedback at several scales
-to check assumptions and stay on course.
+The purpose of feedback is to guide lesson development
+so that authors aren't designing and arguing in a vacuum.
+What people immersed in the lessons think needs fixing
+can often differ from what learners think.
 
 Micro-scale feedback can be gathered by an instructor while teaching a particular lesson.
-Learners can provide feedback on everything from typographical errors
-and the clarity of quiz questions
-to the order in which topics are presented,
-all of which the instructor should record at the end of each class
-in some shared location (such as a Google Doc or GitHub issues).
+Learners might provide feedback on things like typographical errors,
+clarity/ease of quiz questions or the order in which topics are presented,
+which the instructor can enter into a work-tracking system (e.g., GitHub issues) at the end of class.
 As well as encouraging direct verbal feedback,
 it's a good idea to provide learners with a means to provide feedback anonymously during class
-(e.g., on small pieces of paper like sticky notes or through anonymous surveys).
+(e.g., on small pieces of paper like sticky notes).
 
-Pre- and post-class surveys should be used to uncover larger issues,
-particularly those arising from developers not fully understanding their audience,
-e.g.,
-assuming prior knowledge that learners do not have.
-Post-class surveys are most effective when conducted 30--90 days after class;
-this gives people time to reflect,
-so their feedback will more accurately reflect what they learned
+Pre- and post-class surveys can be used to discover larger macro-scale issues.
+These issues often relate to the fact that it can be difficult for lesson developers
+to fully understand the frame of reference of their audience.
+For instance, a lesson might inadvertently assume prior knowledge that learners do not have,
+which is information that can be collected in a survey.
+If possible, it's a good idea to conduct the post-class survey 30--60 days after the fact.
+This delay allows people time to reflect,
+meaning they are more likely to give accurate feedback on what they learned
 rather than how entertained they were.
 
-\rulemajor{Reduce, re-use, recycle}{rrr}
+Pre- and post-class surveys also are essential in focusing in on the audience for a given lesson.
+Referring back to Rule~1,
+it can sometimes be difficult for lesson developers to understand the frame of a given learner,
+so surveys (particularly post-class surveys) can reveal hidden assumed knowledge
+that can be expanded on or acknowledged to refine the target audience.
+This is one example of a larger guideline:
+when processing feedback,
+separate items that apply to specific lessons (and can guide its future development)
+from those that are better tackled elsewhere
+(e.g., by developing new lessons or through organizational change).
 
+\rulemajor{Reduce, re-use, recycle}
+
+Do not create a new lesson if there is an existing one that you can use or contribute to.
 Just as a scholar would not write a paper without a literature review,
-an instructor should not create a new lesson if there is an existing one they could use or contribute to.
-A short online search can discover if someone has written what you need,
-whether it is complementary to your goals,
-and if it can be tweaked or modified to meet your needs.
+the same holds for lessons.
+Before writing that introduction to the Bash command line,
+for example,
+do a search:
+has anybody else written it?
+Is it complementary to your goals?
+Could it be tweaked or modified to meet your own goals?
+Could your planned lesson be tweaked to compliment the existing lessons so that topics aren't duplicated?
 
 Before re-using content,
-make sure to check its license.
+make sure to check that the lesson is licensed in an open manner.
 Both Programming Historian and the Carpentry projects
-use the Creative Commons--Attribution license
+use the Creative Commons - Attribution license
 (\url{https://creativecommons.org/licenses/by/4.0/}),
@@
 as long as they cite the original source.
-Other Creative Commons licenses may restrict commercial use
-and/or creation of derivative materials.
 
-The question of licensing also arises when recycling lesson components
+The same questions of re-use come when thinking about recycling content of a lesson,
 such as images, data, figure, or code.
-If the license does not cover them explicitly,
-ask permission as you would for any other academic material.
+Does the license cover that as well?
+If not,
+then ask permission,
+just as you would for any other material.
 
-The converse of this rule is to make the license for your lessons explicit and discoverable.
+The converse of this rule is to license your own lessons in a similar open manner,
+and to make them discoverable.
 For example,
-when lessons are published (\ruleref{publish}),
-make sure that that keywords such as ``CC-BY'' appear in their bibliography entries
+when lessons are published (Rule~6),
+make sure that they have the usual keywords in their bibliographic entries
 and HTML page headers.
 
-\rulemajor{Link to other resources}{link}
+\rulemajor{Link to other resources}
 
-Learners are unlikely to absorb everything they need to know about a topic from your lesson alone.
+Most learners are unlikely to absorb everything there is to know about a topic
+from your lesson alone.
 This is partly a matter of scope---any interesting subject is too large
@@
 As Caulfield has argued \cite{choral-explanations},
-the best way to use the Internet is to provide a chorus of explanations
+the best resources provide a chorus of explanations
 that offer many angles and approaches for any given topic,
-each of which may be the best fit for a different set of needs.
+each of which may be the best fit for a different subset of people.
 
-Collaboratively-developed lessons should direct learners to these resources at strategic points.
-Textbooks,
+Find these resources and direct learners to them at strategic points.
+These resources may include textbooks,
 technical documentation,
 videos,
-web pages,
-threads on Quora or mailing lists:
+and web pages;
 if a community or discussion forum exists for the topic,
-it is worth including.
+be sure to include them as well.
 
@@
 and maintaining it is even more so,
-which makes building community around lessons (\ruleref{community}) all the more important.
+which makes building community around lessons (Rule~2) all the more important.
 In particular,
@@
 they should both be able to propose updates, corrections, and additions to lessons,
-and know that they are encouraged to do so (\ruleref{empower}).
+and know that they are encouraged to do so (Rule~5).
 
-\rulemajor{You can't please everyone}{everyone}
+\rulemajor{You can't please everyone}
 
@@
 Similarly,
-lessons on ecology for learners in Utah and Vietnam
+lessons on ecology for learners in Utah and Florida
 will probably be more relatable if they use different examples.
-A community may therefore maintain several differently-oriented or differently-paced lessons
-on a single topic,
-just as programming languages provide several different libraries for doing the same general thing
-with different levels of performance and complexity.
-
-Similarly,
+Equally,
 no lesson development community can serve all purposes.
@@
 while others may prefer a ``measure twice, cut once'' approach.
+
 If there are complementary ways to explain something,
@@
 even if contributors \emph{do not} disagree:
-weighing alternatives fosters higher-order thinking.
+weighing alternatives fosters higher-order thinking,
+and as Caulfield has written,
+the most popular online question and answer sites
+are successful in part because they present a chorus of explanations
+geared at different levels and needs
+rather than a single one \cite{choral-explanations}.
 
@@
 \cite{producing-oss}.
-As discussed in \ruleref{empower},
+As discussed in Rule~5,
 the first step is to have a clear governance structure and a clear, permissive license.
 Minor disagreements should be discussed openly and respectfully.
-If they turn out not to be so minor after all,
+If they cannot be resolved---i.e., if they turn out not to be so minor after all---then
 contributors should split off and evolve the lesson in the way they see best.
@@
 % step-by-step instructions.
-
-\newpage
 
@@
 
-\newpage
-
-%%%%% Figure  dissociation
-\begin{figure}[ht]  % h = here, t = top, b = bottom, p = float
-\includegraphics[width=\linewidth]{figure1}
-\caption{Graphical abstract of 10 simple rules for collaborative lesson development}
-\label{figure1}
-\end{figure}
-%
-
-\begin{mdframed}
-
-\noindent
-Box 1: Learner Profile
-
-Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
-While fluent in both Spanish and English,
-he has a hearing disability that sometimes makes it hard for him to understand lectures,
-particularly in noisy environments.
-Other than using Excel, Word, and the Internet,
-Jorge's most significant previous experience with computers is
-helping his sister build a WordPress site for the family business.
-
-Jorge needs to measure properties of soil from nearby farms
-using a handheld device that sends text files to his computer.
-Right now, Jorge has to open each file in Excel,
-crop the first and last points,
-and calculate an average.
-This workshop will show Jorge how to write a little Python program
-to read the data,
-select the right values from each file,
-and calculate the required statistics.
-
-\end{mdframed}
-
 \end{document}

### 2017-07-04 paper.tex
Tentative improvements of fig1
- make it included as pdf so it is not pixelized/rasterized (generated with inkscape + ghostscript cleaning)
- use a font that is closer (in style and size) to the paper's font
- add some space to breathe
- add crossroad sign (openclipart)
Fixing typos (for #32)

--- 2017-07-01 paper.tex
@@
 Despite the success of openness in software development and the curation of Wikipedia,
-it is an uncommon approach in accademic settings.
+it is an uncommon approach in academic settings.
 Each year,
@@
 and that most researchers complain about how much time teaching takes away from research,
-but the sociology and pscyhology behind this blind spot are out of the scope of this paper.
+but the sociology and psychology behind this blind spot are out of the scope of this paper.
 
@@
 this is not the same as continuous improvement by a large community of contributors.
-The ten simple rules that follow summarize what we have learned about doing that.
+The ten simple rules that follow summarize what we have learned about doing that
+as maintainers, editors, and reviewers of lessons used by tens of thousands of people (Fig. 1, 2).
 
@@
 but can also unintentionally suppress voices.
-Programming Historian makes an ombudsperson is available for private chats and facilitation
+Programming Historian makes an ombudsperson available for private chats and facilitation
 to ensure that no one is excluded.
@@
 \end{figure}
-%
+
+\newpage
+
+%%%%% Figure  dissociation
+\begin{figure}[ht]  % h = here, t = top, b = bottom, p = float
+\includegraphics[width=\linewidth]{figure2}
+\caption{Collaborative open lesson development}
+\label{figure2}
+\end{figure}
+
+\newpage
 

### 2017-07-07 paper.tex
Adding more material on learning objectives.
Closes #31.

@@
 
-Rather than itemizing prior knowledge and learning objectives,
+Instead of starting with learning objectives (\ruleref{practices}),
 it can be helpful to write \emph{learner profiles} to clarify
@@
 these modules can be made more discoverable
-by referencing specific points in the model curricula promulgated by many professional societies.
+by referencing specific points in the model curricula promulgated by many professional societies
+(e.g., as learning objectives).
 Smaller modules are also more approachable for new contributors (\ruleref{empower}).
@@
 
-Pre- and post-class surveys should be used to uncover larger issues,
+Pre- and post-class surveys and interviews should be used to uncover larger issues,
 particularly those arising from developers not fully understanding their audience,
@@
 rather than how entertained they were.
+Clearly-stated learning objectives (\ruleref{practices}) are essential here,
+as they tell assessors what they should be measuring.
 

### 2017-07-07 paper.tex
Adding acknowledgments.
Closes #34.

@@
 as maintainers, editors, and reviewers of lessons used by tens of thousands of people (Fig. 1, 2).
+
+\subsection*{Acknowledgments}
+
+We are grateful to everyone who provided feedback on this paper,
+including
+James Baker,
+Nathan Moore,
+Pariksheet Nanda,
+Tom Pollard,
+Byron Smith,
+and Andrew Walker.
+We are also grateful to the hundreds of people who have contributed to
+Programming Historian, Data Carpentry, Software Carpentry, and Library Carpentry
+over many years.
 

### 2017-07-07 paper.tex
Commenting on file formats being an extra obstacle.
Closes #36.

@@
 which has a notoriously steep learning curve \cite{git-survey}.
-Google Docs and wikis do not support pre-merge review,
+
+Complicating matters further,
+some file formats make collaboration easier or more difficult.
+Despite their ubiquity,
+open source version control systems do not directly support review or merge
+of Microsoft Office or LibreOffice file formats,
+which raises an additional burden for newcomers \cite{jacobs}.
+Programmers may look down on Google Docs and wikis
+for their lack of pre-merge review and other capabilities,
 but their low barrier to entry make them more welcoming to newcomers.
@@
 
+\bibitem{jacobs}
+Jacobs C.~T., Gorman G.~J., Rees H.~E., and Craig L.~E.:
+Experiences With Efficient Methodologies for Teaching Computer Programming to Geoscientists
+\newblock Journal of Geoscience Education
+\newblock 64(3),
+\newblock 2016,
+\newblock https://doi.org/10.5408/15-101.1
+
 \end{thebibliography}

### 2017-07-07 paper.tex
Formatting fixes to get it through arxiv.org

@@
 % Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
+% \usepackage{nameref,hyperref}
 
@@
 \setlength{\headheight}{27.023pt}
-\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+% \lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
 \rfoot{\thepage/\pageref{LastPage}}
@@
 \fancyfootoffset[L]{2.25in}
-\lfoot{\sf PLOS}
+% \lfoot{\sf PLOS}
 
@@
 Several open frameworks are available to facilitate development of new lessons,
-such as learnr (\url{https://rstudio.github.io/learnr}),
-Morea (\url{https://morea-framework.github.io}),
-and DataCamp's templates (\url{https://www.datacamp.com/teach/documentation}).
+such as learnr (\texttt{https://rstudio.github.io/learnr}),
+Morea (\texttt{https://morea-framework.github.io}),
+and DataCamp's templates (\texttt{https://www.datacamp.com/teach/documentation}).
 
@@
 it is important to publish lessons in ways that traditional academic systems can digest.
-One is to give releases DOIs supplied by providers such as Zenodo (\url{https://zenodo.org/})
-or DataCite (\url{https://www.datacite.org/}).
+One is to give releases DOIs supplied by providers such as Zenodo (\texttt{https://zenodo.org/})
+or DataCite (\texttt{https://www.datacite.org/}).
 Contributors can be listed as authors
@@
 Each time the lesson is published,
-names (and identifiers such as ORCIDs (\url{https://orcid.org}))
+names (and identifiers such as ORCIDs (\texttt{https://orcid.org}))
 should be gathered for all contributors.
@@
 use the Creative Commons--Attribution license
-(\url{https://creativecommons.org/licenses/by/4.0/}),
+(\texttt{https://creativecommons.org/licenses/by/4.0/}),
 which allows people to share and adapt material for any purpose
@@
 \newblock Software Carpentry: Instructor Training; 2016.
-\newblock \url{https://zenodo.org/record/57571#.WS8huDOZPdQ}.
+\newblock \texttt{https://zenodo.org/record/57571\#.WS8huDOZPdQ}.
 
@@
 Software Carpentry: The Unix Shell; 2015.
-\newblock \url{https://zenodo.org/record/27355#.WS8lajOZPdQ}.
+\newblock \texttt{https://zenodo.org/record/27355\#.WS8lajOZPdQ}.
 
@@
 Software Carpentry: The Unix Shell; 2017.
-\newblock \url{https://zenodo.org/record/278226#.WS74tTOZPdQ}.
+\newblock \texttt{https://zenodo.org/record/278226\#.WS74tTOZPdQ}.
 
@@
 Caulfield M. Choral Explanations; 2016.
-\newblock \url{https://hapgood.us/2016/05/13/choral-explanations/}.
+\newblock \texttt{https://hapgood.us/2016/05/13/choral-explanations/}.
 
@@
 \newblock Global Developer Survey.
-\newblock \url{http://get.gitlab.com/global-developer-survey/}, 2016.
+\newblock \texttt{http://get.gitlab.com/global-developer-survey/}, 2016.
 

### 2017-09-27 paper.tex
Fixing DOI
Changing Greg's affiliation
Handling some of the smaller tasks as per #40 See checklist, not entirely sure about tweaks to Rule #3 and #5

--- 2017-07-07 paper.tex
@@
 \textbf{6} University of Waterloo / i2millig@uwaterloo.ca \\
-\textbf{7} Rangle.io / gvwilson@third-bit.com \\
+\textbf{7} DataCamp / gvwilson@third-bit.com \\
 
@@
 Despite the success of openness in software development and the curation of Wikipedia,
-it is an uncommon approach in academic settings.
+it is an uncommon approach in the academic instructional setting.
 Each year,
@@
 they develop and maintain their course materials in isolation.
-This is curious given that research depends critically on sharing,
-and that most researchers complain about how much time teaching takes away from research,
+Given that academic research often depends on sharing, the differing
+approach to developing pedagogical materials is interesting,
 but the sociology and psychology behind this blind spot are out of the scope of this paper.
@@
 
-\subsection*{Acknowledgments}
-
-We are grateful to everyone who provided feedback on this paper,
-including
-James Baker,
-Nathan Moore,
-Pariksheet Nanda,
-Tom Pollard,
-Byron Smith,
-and Andrew Walker.
-We are also grateful to the hundreds of people who have contributed to
-Programming Historian, Data Carpentry, Software Carpentry, and Library Carpentry
-over many years.
-
 \rulemajor{Clarify your audience}{audience}
@@
 The first requirement for building lessons together is
-to know who they are being built for.
+to know whom they are being built for.
 ``Archaeology students'' is far too vague:
@@
 first-year students who need an introduction to the field,
-graduate students who intend to specialize in the sub-discipline which is the lesson's focus,
+graduate students who intend to specialize in the sub-discipline that is the lesson's focus,
 or someone in between?
@@
 
-Decades of pedagogical research has yielded many insights into
+Decades of pedagogical research have yielded many insights into
 how best to build and deliver lessons \cite{hlw}.
 Unfortunately,
-many college and university faculty have little or no training in education \cite{brownell},
+many college and university faculty have little or no formal training in education \cite{brownell},
 so this knowledge is rarely applied in the classroom.
@@
 
-An example of how to teach these practices is Software Carpentry's instructor training program.
+An example of how to teach these practices is Software Carpentry's (\url{https://software-carpentry.org/}) instructor training program.
 First offered in 2012,
@@
 which raises an additional burden for newcomers \cite{jacobs}.
-Programmers may look down on Google Docs and wikis
-for their lack of pre-merge review and other capabilities,
-but their low barrier to entry make them more welcoming to newcomers.
+While Google Docs and wikis lack some capabilities, such as
+full-fledged pre-merge review (although ``suggest mode'' 
+mitigates this to some degree), 
+their low barrier to entry makes them more welcoming to newcomers.
 
@@
 or asking them to add new exercises and tweak diagrams and references
-can all provide an on-ramp for people who might question their authority or ability to change the main body of a lesson.
+can all provide an on-ramp for people who might question their own authority or ability to change the main body of a lesson.
 Equally,
@@
 gives new contributors an early reward for taking part. 
+It can also help with the formal and informal recognition
+of labour, discussed in the following rule.
 
@@
 and archive old versions in a discoverable location.
-Also make sure that everyone involved knows what ``done'' look like,
+Also make sure that everyone involved knows what ``done'' looks like,
 i.e.,
 which outstanding issues have to be addressed
-and how it has to be formatted
+and how they have to be formatted
 in order for the next release to go out.
@@
 adopting the model of collaborative software development
-to make more robust and sustainable lessons
+to make more robust and sustainable lessons in all domains
 that can be continually improved by those who use them.
@@
 with more impact and less effort.
+
+\subsection*{Acknowledgments}
+
+We are grateful to everyone who provided feedback on this paper,
+including
+James Baker,
+Nathan Moore,
+Pariksheet Nanda,
+Tom Pollard,
+Byron Smith,
+and Andrew Walker.
+We are also grateful to the hundreds of people who have contributed to
+Programming Historian, Data Carpentry, Software Carpentry, and Library Carpentry
+over many years.
 
@@
 \newblock 2016,
-\newblock https://doi.org/10.5408/15-101.1
+\newblock 10.5408/15-101.1
 

### 2017-10-03 paper.tex
Reverted a sentence I put into rule 5 I don't think we need signposting in such proximity.
Added framing

--- 2017-07-07 paper.tex
@@
 and that most researchers complain about how much time teaching takes away from research,
-but the sociology and psychology behind this blind spot are out of the scope of this paper.
+but the sociology and psychology behind this blind spot are beyond the scope of this paper.
 
 The authors have many years of experience with community-developed lessons
-in the context of research computing in the sciences and humanities
+in the context of research computing in the sciences and humanities,
 through organizations like Software Carpentry and Programming Historian.
@@
 
-All open education projects satisfy the first criterion by definition,
-but few satisfy the second.
-While their lessons are occasionally updated by a small team
+All open education projects (e.g. massive open online courses) satisfy the first criterion by definition,
+but very few satisfy the second.
+In other words,
+while it is common for open education projects to be occasionally updated by an individual or small team
 (as happens when a new edition of a book is edited and published),
@@
 as maintainers, editors, and reviewers of lessons used by tens of thousands of people (Fig. 1, 2).
+By following these rules,
+we contend that it is possible to create higher quality lessons 
+than could be created by an individual or small team,
+both in terms of accuracy and pedagogy.
+As an added bonus, the lessons are always up-to-date 
+and require less time per author to develop and maintain.
 
@@
 First offered in 2012,
-is now a two-day course delivered both in-person and online
+it is now a two-day course delivered both in-person and online
 \cite{lessons-learned,instructor-training,how-to-teach-programming}.

### 2017-10-09 paper.tex

--- 2017-09-27 paper.tex
@@
 they develop and maintain their course materials in isolation.
+
 Given that academic research often depends on sharing, the differing
 approach to developing pedagogical materials is interesting,
-but the sociology and psychology behind this blind spot are out of the scope of this paper.
+but the sociology and psychology behind this blind spot are beyond the scope of this paper.
 
 The authors have many years of experience with community-developed lessons
-in the context of research computing in the sciences and humanities
+in the context of research computing in the sciences and humanities,
 through organizations like Software Carpentry and Programming Historian.
@@
 gives new contributors an early reward for taking part.
-It can also help with the formal and informal recognition
-of labour, discussed in the following rule.
 

### 2017-10-10 paper.tex
figure legend and image update

@@
 and require less time per author to develop and maintain.
+
+\textbf{Fig 1.
+Graphical abstract of ten simple rules for collaborative lesson development.} \textbf{1.} To clarify your audience, consider writing learner profiles (see Box 1).
+\textbf{2.} Make lessons modular by breaking them into small, single-purpose modules.
+\textbf{3.} Teach your instructors the best practices for developing, delivering, and maintaining lessons.
+\textbf{4.} Encourage and empower contributors by making the contribution process transparent and straightforward.
+\textbf{5} Build a community around lessons by creating opportunities for participation and mentorship.
+\textbf{6.} Publish new version periodically and recognize contributors by their unique identifiers (e.g., ORCID).
+\textbf{7.} Evaluate lessons during and after class for a complete picture of the lessons' efficacy.
+\textbf{8.} Reduce, re-use, or recycle lessons before creating a new one from scratch.
+\textbf{9.} Link to other resources that complement the lesson content.
+\textbf{10.} Remember that you can't please everyone in your audience or community.
 

### 2017-10-10 paper.tex

--- 2017-10-09 paper.tex
@@
 
+Lessons take significant effort to build and maintain.
 The collaborative development methods pioneered by the open source software community
-offer a way to create lessons that are open, accessible, and sustainable.
-This paper presents ten simple rules for doing this
-drawn from our experience with several successful projects.
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step.
-% Author Summary not valid for PLOS ONE submissions.
-\section*{Author summary}
-
-Lessons take significant effort to build and  maintain.
-We have found that
-the collaborative development methods pioneered by the open source software community
-are an effective, economical way to create and sustain lessons
+offer an effective, economical way to create and sustain lessons
 that can be used by large numbers of people in a wide variety of contexts.
 The ten simple rules outlined in this paper
-summarize the best practices that we have helped evolve
-in several successful open education projects aimed at researchers and research librarians
+summarize the best practices that have been implemented
+by several successful open education projects aimed at researchers and research librarians
 in a wide range of disciplines.

### 2017-10-16 reviewer-response.md
First draft of a reviewer response letter Mostly line-by-line responses as in my field.

Dear Dr. Markel –

Thanks for your engagement with our paper, "Ten simple rules for collaborative lesson development." We appreciated the reviewer's engagement with the piece, and were gratified to see that the ideas generally resonated with them. Please accept our revised version and we look forward to hearing from you soon.

We have provided point-by-point reviewer responses below.

# Reviewer #1

Reviewer #1: The "Ten simple rules for collaborative lesson development" article presents a much needed perspective on teaching materials, and may be just the push needed to move lecturers to move their content online and open such that others may benefit from their materials.

Overall this paper is very well written and to the point. Only a few minor comments:

1) Line 21 - the authors should distinguish the academic setting from the academic teaching setting, since the academic setting does very well in open source software development.

**We have made this explicit.**

2) Lines 24-27 seem to be a personal statement against the academic system.

**We have removed the complaint component, and noted instead the divergence between sharing in academic research and the reluctance to want to do so in a teaching environment.**

3) It would be useful to provide the URL in the text for rule #3 since online best practices for lesson development circles back to online collaborative lesson development.

**We were not crystal clear on this suggestion. We did include a link to Software Carpentry here as we felt that would help people engage with it.**

4) line 141 - question their 'own' authority

**Fixed.**

5) In Rule #5, it would be helpful if the authors redirected to Rule #6 when discussing recognizing contributions.

**On reflection, we liked the divison that we had between Rule #5 and Rule #6, and did not want to rebalance between the two rules. We considered adding some signposting between the two rules, but felt it was unnecessary given their proximity.**

6) Line 251, sustainable lessons 'in all domains'

**Fixed.**

## Reviewer #3

Reviewer #3: The authors of "Ten simple rules for collaborative lesson development" have compiled a very useful set of guidelines for successfully building training materials. I know I am going to use them!

I am very pleased with the contents and the way in which the manuscript highlights a number of very useful tricks to creating useful and valuable teaching material. It was not easy to find an area of the manuscript that could be improved, but I managed to spot a few, minor things and I am hopeful that my comments will be of use to polish up a bit what is already a great compilation of your expertise building these teaching materials.

I find Tip #3 a bit difficult to digest. The flow of the narrative becomes a bit cloudy when it turns into a 5-point lesson within the 10-point document. A reference has been made to a different publication, which should help - but a bit more clarity on how this collection of five points become "best practices" is necessary. Perhaps a paragraph that more clearly states the important content the authors extracted from Wiggins and McTighe and how they used them to develop these "best practices" would be easier to understand.

**Upon reflection, and given the positive feedback from other reviewers, we decided to leave Rule #3 as is.**

In Figure 1.
The images supporting Tip #7 are not very clear to interpret. I would think of something else here. Perhaps an old-style scale instead of 'great' and 'terrible'? Or something similar.

**We changed the figure to help with interpretability. In one sentence, we state the rule and some little extra bit of information that ties the image in the figure to the explanatory information in the body of the manuscript.**

**The figure also has the following modifications:**
- **modified the image for rule #7 to incorporate the above suggestion;**
- **Removed the PLOSlgo reference to our paper.**
- **Made more space around rules 1-4.**
- **Tweaked the icons for rules #3 and #6.**

Regarding Figure 2.
This diagram could be improved with a few arrow movements and slightly different wording. Active voice is always an efficient way to deliver information, so I kindly suggest that the authors could make a bit more fruitful use of it.

For instance, "Lesson materials are outlined and drafted via community input" could be more clearly stated by saying instead: "Outline and draft lesson materials with community input."
"Materials are updated continually" could also be switched into active voice.
"Lesson release..." --> "Release a stable version of the lesson."

Also, the arrow from 'recruit community..." into "Lesson materials..." could go and meet the arrow going from 'an individual or group...' towards 'Lesson materials...' It adds clarity.

**Figure 2 was updated to include boxes labeled in active voice and arrows between boxes that better represent connections among concepts. We also added a white background and added additional file formats.**

Finally, just a minor comment (not a request for review) that at first, the thought of inner references from one 'Tip' to the other was worrisome to me, but the authors were able to disperse those fears by the end of the article. I actually quite like the internal references to other points in the list. It is very useful!

That's all. Thanks for sharing this insight with the community. As I mentioned before, I am certain these Tips will be very useful to many!

## Reviewer #4

Reviewer #4: The authors present ten rules for successfully developing teaching materials in a collaborative, open framework. The idea of collaborative and open lesson development is an important and relevant one, and so in that sense this manuscript would be a useful contribution to the literature. The rules are generally well-chosen, and fit together well.

My major criticism of the paper is that it is overly sparse in its framing. This is especially true of the beginning (the abstract is currently too short to be very useful), but also at the end. I think that the authors could do a better job of 1) motivating the need for this new approach to instructional material design and maintenance and 2) addressing criticisms that more conventional practitioners might have. These concerns could include reservations about the accuracy of open lessons, the coherency of modules made by contributors that likely span a range of experience in a given discipline, the reluctance of academic administrators to have their students taught with course materials developed by third parties, etc.

We agreed with this criticism and revised the paper accordingly. We did this in two distinct ways.

**First, in the submission version, we had both an abstract and an author summary right after each other. We removed the author summary and worked the text into the abstract, fleshing things out. Given that our paper is not technical we did not believe that an author summary was warranted.**

**Secondly, we expanded the introduction itself. We did so by clarifying the criteria around open education projects, and further reinforced the benefits of this approach: noting higher-qualtiy lessons, ensuring that lessons were up-to-date, and how individual time on behalf of authors is less. Some of this was in the lessons but we agreed that the introdution needed to frame.**

**We did not address concerns around how some might have reservations about "the coherency of modules made by contributors that likely span a range of experience" as we felt that coherence can come from Rule #3.**

**We did not address "the reluctance of academic administrators to have their students taught with course materials developed by third parties," because the paper is short, sharp and concise and generally avoids getting sidetracked on peripheral issues such as this.**

I recognize that there are word limits on manuscript of this kind, but I think that with further revision the writing could be tightened up a bit to allow for this additional framing and discussion. The authors have a chance here not only to show how to do this sort of lesson development, but also motivate people to consider doing it in the first place. Not everyone is ready to embrace this new way of producing pedagogical materials, but perhaps they would be if a stronger case were made. I guess all that is to say I'd prefer a bit more space spent on the 'why' instead of just on the 'how'.

I also think that the manuscript would be improved with another few passes to smooth out the text. There are a number of spots where transitions are a bit awkward. I also noticed a number of grammatical and proofreading errors that would need to be corrected before publication. I've pointed out some but not all of them below.

**We have done a pass, incorporating the flagged changes as well as other minor ones. Our sincere thanks for the close read.**

Specific comments:

Acknowledgements usually go at the end, not before the 10 rules?

**Fixed.**

Aside from their parenthetical mention at the end of the Introduction, neither of the figures is referred to in the main set of rules. I am not sure if both of these are necessary. Figure two, especially, encompasses a number of rather complex ideas about community building and lesson development that are not directly discussed at any length within the text. The second figure seems the more useful of the two, but in any event, I think both warrant more attention in the text or in captions if they are to be kept.

**Our intention for figure 1 was to provide a graphical abstract or a concise summary captures the content of the article in a single glance. We think it serves its purpose well. Similarly, Figure 2 covers the main rules and their interplay. Both figures are now explicitly referenced in the paper. We also provided in-depth captions for the figures which more strongly tie them into the paper.**

pg 2, line 50: who -> whom

**Fixed.**

pg 2, line 53: which -> that

**Fixed.**

pg 3, line 77: has -> have

**Fixed.**

pg 3, line 78: I'd say 'formal training' here. There are many ways to learn to do something. Without this adjustment you risk coming off as dismissive or even derogatory towards faculty, which I don't think is your intent.

**Agreed - we have made this change.**

pg 4, line 117: I'd also revise this bit about programmers 'looking down' on Google Docs and the like. It contributes to the notion that programmers think they are better than others, which is not something that belongs in a manuscript about inclusive lesson development. Recent versions of Google Docs also allow for 'suggesting' mode edits, which could be considered a version of 'pre-merge review', in that the changes can be discussed, accepted, or rejected after they are initially made/proposed.

**Agreed – we have revised this to note both suggest mode but also that the low barrier to entry for Google Docs can make it an attractive starting point.**

pg 5, line 168: look -> looks

**Fixed.**

pg 5, line 168: it has -> they have

**Fixed.**

### 2017-10-16 reviewer-response.md
fixing minor typos in reviewer-response.md

@@
 
-**On reflection, we liked the divison that we had between Rule #5 and Rule #6, and did not want to rebalance between the two rules.
+**On reflection, we liked the division that we had between Rule #5 and Rule #6, and did not want to re-balance between the two rules.
 We considered adding some signposting between the two rules, but felt it was unnecessary given their proximity.**
@@
 **Secondly, we expanded the introduction itself.
-We did so by clarifying the criteria around open education projects, and further reinforced the benefits of this approach: noting higher-qualtiy lessons, ensuring that lessons were up-to-date, and how individual time on behalf of authors is less.
-Some of this was in the lessons but we agreed that the introdution needed to frame.**
+We did so by clarifying the criteria around open education projects, and further reinforced the benefits of this approach: noting higher-quality lessons, ensuring that lessons were up-to-date, and how individual time on behalf of authors is less.
+Some of this was in the lessons but we agreed that the introduction needed to frame.**
 

### 2017-10-16 paper.tex

@@
 
+\textbf{Fig 2. Collaborative open lesson development.} Following the decision to develop 
+lessons, activities focus on lesson development as well as community building. 
+Boxes surrounded by dotten lines represent community contributions to lessons.
+Dashed arrows represent connections to activities outside the original lesson design.
+The box enclosed in a dash-dotted line represents unaffiliated learning resources.
+
 \rulemajor{Clarify your audience}{audience}
@@
 Finally,
-teaching people how to teach is a great way to introduce them to each other and build community.
+teaching people how to teach is a great way to introduce them to each other and build community (Fig. 2).
 
@@
 specific versions of lessons should be published or released periodically
-so that learners or instructors have something stable to refer to for the duration of their use.
+so that learners or instructors have something stable to refer to for the duration of their use (Fig. 2).
 Periodic releases also provide an opportunity
@@
 It is therefore critical to gather and act on feedback at several scales
-to check assumptions and stay on course.
+to check assumptions and stay on course (Fig. 2).
 
@@
 that offer many angles and approaches for any given topic,
-each of which may be the best fit for a different set of needs.
+each of which may be the best fit for a different set of needs (Fig. 2).
 

### 2017-10-23 paper.tex

@@
 
-One popular lesson development methodology is presented in \cite{wiggins-mctighe}.
-When this is used,
-lessons are built by:
-
-\begin{enumerate}
-
-  \item
-    identifying learning objectives,
-
-  \item
-    creating \emph{summative assessments} to determine whether those objectives have been met,
-
-  \item
-    designing \emph{formative assessments} to gauge learners' progress
-    and give them a chance to practice key skills,
-
-  \item
-    putting those formative assessments in order,
-
-  \item
-    and only then writing lessons to connect each to the next.
-
-\end{enumerate}
-
-\noindent
+By way of example, Software Carpentry encourages its volunteers to use
+the popular lesson development methodology presented by \cite{wiggins-mctighe},
+whereby learning objectives and assessments are created \emph{before} any lesson materials are developed.
+In particular, \emph{summative assessments} are created to determine whether the learning objectives have been met,
+and \emph{formative assessments} are created to gauge learners' progress and to give them a chance to practice key skills.
+The completed formative assessments are put in order and only then are the lessons written, 
+with the aim of connecting each formative assessment to the next.
 This method is effective in its own right,
@@
 
-An example of how to teach these practices is Software Carpentry's (\url{https://software-carpentry.org/}) instructor training program.
+An example of how to teach such pedagogical practices
+is Software Carpentry's (\url{https://software-carpentry.org/}) instructor training program.
 First offered in 2012,
@@
 \cite{lessons-learned,instructor-training,how-to-teach-programming}.
-This program teaches good pedagogical practices,
-and introduces everyone who takes it to who Software Carpentry's lessons are for,
+In addition to a focus on pedagogy,
+the course introduces everyone who takes it to who Software Carpentry's lessons are for,
 how they are delivered,

### 2017-10-24 paper.tex
Removed a random URL
added paragraph on mentoring
related to https://github.com/swcarpentry/collaborative-lesson-development/issues/56

@@
 
+In 2015, Software Carpentry established a 'Mentoring Subcommittee' to support instructors as they progressed through training, teaching, curriculum development.
+The Mentoring Subcommittee has promoted community-building by providing virtual spaces where instructors from all over the world can share success stories and discuss strategies for overcoming challenges.
+This sharing of expertise and experience between geographically separated instructors has helped strengthen the community around lessons and provides insight into how lessons can be improved (see rule #7).
+
 Finally,

### 2017-10-24 paper.tex
mention code of conduct

--- 2017-10-23 paper.tex
@@
 to ensure that no one is excluded.
+Software Carpentry operates by a Code of Conduct that outlines acceptable standards of behavior for our community members and those interacting with the Carpentries at in-person events and in virtual spaces.
+Community members on a 'Policy Subcommittee' serve as advocates for the CoC and adjudicate reported Code of Conduct violations.
+Together, the Code of Conduct and the Policy Subcommittee help ensure that all members of the community are treated with respect.
 

### 2017-10-29 paper.tex
CoC

--- 2017-10-23 paper.tex
@@
 to ensure that no one is excluded.
+Software Carpentry operates by a Code of Conduct that outlines acceptable standards of behavior for our community members
+and those interacting with the Carpentries at in-person events and in virtual spaces.
+Community members on a Policy Subcommittee serve as advocates for the Code of Conduct
+and adjudicate reported violations.
 

### 2017-10-29 paper.tex

--- 2017-10-24 paper.tex
@@
 An example of how to teach such pedagogical practices
-is Software Carpentry's (\url{https://software-carpentry.org/}) instructor training program.
+is Software Carpentry's instructor training program.
 First offered in 2012,

### 2017-10-29 paper.tex
Final formatting

--- 2017-10-29 paper.tex
@@
 
+\begin{mdframed}
+\noindent
 \textbf{Fig 1.
@@
 \textbf{10.} Remember that you can't please everyone in your audience or community.
+\end{mdframed}
 
+\begin{mdframed}
+\noindent
 \textbf{Fig 2. Collaborative open lesson development.} Following the decision to develop 
@@
 The box enclosed in a dash-dotted line represents unaffiliated learning resources.
+\end{mdframed}
 
@@
 gives new contributors an early reward for taking part.
+
+In 2015,
+Software Carpentry established a Mentoring Subcommittee to support instructors
+as they progress through training, teaching, curriculum development.
+The Mentoring Subcommittee has promoted community building
+by providing virtual spaces where instructors from all over the world can share success stories
+and discuss strategies for overcoming challenges.
+This has helped strengthen the community and provided insight into how lessons can be improved
+(\ruleref{evaluate}).
 

### 2018-01-12 paper.tex
Modified affiliations section

@@
 \bigskip
-\textbf{1} Douglas Mental Health University Institute, McGill University / gdevenyi@gmail.com \\
-\textbf{2} Univ Lyon, UJM-Saint-Etienne, F-42023, France / remi.emonet@univ-st-etienne.fr \\
-\textbf{3} The University of Texas at Austin / rayna.harris@utexas.edu \\
-\textbf{4} The University of Texas at Tyler / khertweck@uttyler.edu \\
-\textbf{5} CSIRO Oceans and Atmosphere / irving.damien@gmail.com \\
-\textbf{6} University of Waterloo / i2millig@uwaterloo.ca \\
-\textbf{7} DataCamp / gvwilson@third-bit.com \\
+\textbf{1} Douglas Mental Health University Institute, McGill University, Montreal, Quebec, Canada \\
+\textbf{2} Univ Lyon, UJM-Saint-Etienne, Saint-\'{E}tienne, France \\
+\textbf{3} The University of Texas at Austin, Austin, Texas, USA \\
+\textbf{4} The University of Texas at Tyler, Tyler, Texas, USA \\
+\textbf{5} CSIRO Oceans and Atmosphere, Hobart, Tasmania, Australia \\
+\textbf{6} University of Waterloo, Waterloo, Ontario Canada \\
+\textbf{7} DataCamp, Toronto, Ontario, Canada, gvwilson@third-bit.com \\
 

### 2018-01-14 paper.tex
Rearranging bibliography entries so that citations appear in numerical order
Moved "box 1" to right after rule 1, as requested
Fixed minor typo; added SC & PH URLs (Happy to be reverted.. maybe we removed URLs for reason)

@@
 in the context of research computing in the sciences and humanities,
-through organizations like Software Carpentry and Programming Historian.
+through organizations like Software Carpentry (\url{https://software-carpentry.org/}) and Programming Historian (\url{https://programminghistorian.org/}).
 Software Carpentry was founded in 1998 to teach scientists basic computing skills,
@@
 lessons, activities focus on lesson development as well as community building. 
-Boxes surrounded by dotten lines represent community contributions to lessons.
+Boxes surrounded by dotted lines represent community contributions to lessons.
 Dashed arrows represent connections to activities outside the original lesson design.
@@
 and a typical learner profile is presented in Box~1.
+
+\begin{mdframed}
+
+\noindent
+Box 1: Learner Profile
+
+Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
+While fluent in both Spanish and English,
+he has a hearing disability that sometimes makes it hard for him to understand lectures,
+particularly in noisy environments.
+Other than using Excel, Word, and the Internet,
+Jorge's most significant previous experience with computers is
+helping his sister build a WordPress site for the family business.
+
+Jorge needs to measure properties of soil from nearby farms
+using a handheld device that sends text files to his computer.
+Right now, Jorge has to open each file in Excel,
+crop the first and last points,
+and calculate an average.
+This workshop will show Jorge how to write a little Python program
+to read the data,
+select the right values from each file,
+and calculate the required statistics.
+
+\end{mdframed}
 
@@
 
+\bibitem{brownell}
+Brownell S.~E. and Tanner K.~D.:
+Barriers to faculty pedagogical change: Lack of training, time, incentives, and...tensions with professional identity?
+\newblock CBE Life Sciences Education,
+\newblock 11(4), 2012,
+\newblock 10.1187/cbe.12-09-0163.
+
 \bibitem{wiggins-mctighe}
@@
 
+\bibitem{git-survey}
+GitLab
+\newblock Global Developer Survey.
+\newblock \url{http://get.gitlab.com/global-developer-survey/}, 2016.
+
+\bibitem{jacobs}
+Jacobs C.~T., Gorman G.~J., Rees H.~E., and Craig L.~E.:
+Experiences With Efficient Methodologies for Teaching Computer Programming to Geoscientists
+\newblock Journal of Geoscience Education
+\newblock 64(3),
+\newblock 2016,
+\newblock 10.5408/15-101.1
+
 \bibitem{shell2015}
@@
 \newblock O'Reilly; 2005.
-
-\bibitem{git-survey}
-GitLab
-\newblock Global Developer Survey.
-\newblock \url{http://get.gitlab.com/global-developer-survey/}, 2016.
-
-\bibitem{brownell}
-Brownell S.~E. and Tanner K.~D.:
-Barriers to faculty pedagogical change: Lack of training, time, incentives, and...tensions with professional identity?
-\newblock CBE Life Sciences Education,
-\newblock 11(4), 2012,
-\newblock 10.1187/cbe.12-09-0163.
-
-\bibitem{jacobs}
-Jacobs C.~T., Gorman G.~J., Rees H.~E., and Craig L.~E.:
-Experiences With Efficient Methodologies for Teaching Computer Programming to Geoscientists
-\newblock Journal of Geoscience Education
-\newblock 64(3),
-\newblock 2016,
-\newblock 10.5408/15-101.1
 
@@
 
-\newpage
-
-\begin{mdframed}
-
-\noindent
-Box 1: Learner Profile
-
-Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
-While fluent in both Spanish and English,
-he has a hearing disability that sometimes makes it hard for him to understand lectures,
-particularly in noisy environments.
-Other than using Excel, Word, and the Internet,
-Jorge's most significant previous experience with computers is
-helping his sister build a WordPress site for the family business.
-
-Jorge needs to measure properties of soil from nearby farms
-using a handheld device that sends text files to his computer.
-Right now, Jorge has to open each file in Excel,
-crop the first and last points,
-and calculate an average.
-This workshop will show Jorge how to write a little Python program
-to read the data,
-select the right values from each file,
-and calculate the required statistics.
-
-\end{mdframed}
-
 \end{document}

### 2018-01-14 paper.tex
Added programming historian citation (again happy to be rolled back!)

@@
 in the context of research computing in the sciences and humanities,
-through organizations like Software Carpentry (\url{https://software-carpentry.org/}) and Programming Historian (\url{https://programminghistorian.org/}).
+through organizations like Software Carpentry and Programming Historian \cite{proghist}.
 Software Carpentry was founded in 1998 to teach scientists basic computing skills,
@@
 
+\bibitem{proghist}
+Afanador-Llach MJ, Baker J, Crymble A, Gayol V, Gibbs F, Lincoln M, McDaniel C, Milligan I, Parr J, Castro AR, Knuppel AS, Visconti A, Walsh B, Wieringa J.
+\newblock Programming Historian,
+\newblock \url{https://programminghistorian.org/}.
+
 \bibitem{hlw}

### 2018-02-07 highlighted.tex

--- 2018-01-14 paper.tex
@@
 
+% Highlight changes.
+\usepackage{soul}
+
 % Use the PLoS provided BiBTeX style
@@
 \bigskip
-\textbf{1} Douglas Mental Health University Institute, McGill University, Montreal, Quebec, Canada \\
-\textbf{2} Univ Lyon, UJM-Saint-Etienne, Saint-\'{E}tienne, France \\
-\textbf{3} The University of Texas at Austin, Austin, Texas, USA \\
-\textbf{4} The University of Texas at Tyler, Tyler, Texas, USA \\
-\textbf{5} CSIRO Oceans and Atmosphere, Hobart, Tasmania, Australia \\
-\textbf{6} University of Waterloo, Waterloo, Ontario Canada \\
-\textbf{7} DataCamp, Toronto, Ontario, Canada, gvwilson@third-bit.com \\
+\textbf{1} \hl{Douglas Mental Health University Institute, McGill University, Montreal, Quebec, Canada} \\
+\textbf{2} \hl{Univ Lyon, UJM-Saint-Etienne, Saint-\'{E}tienne, France} \\
+\textbf{3} \hl{The University of Texas at Austin, Austin, Texas, USA} \\
+\textbf{4} \hl{The University of Texas at Tyler, Tyler, Texas, USA} \\
+\textbf{5} \hl{CSIRO Oceans and Atmosphere, Hobart, Tasmania, Australia} \\
+\textbf{6} \hl{University of Waterloo, Waterloo, Ontario Canada} \\
+\textbf{7} \hl{DataCamp, Toronto, Ontario, Canada, gvwilson@third-bit.com} \\
 
@@
 \noindent
-Box 1: Learner Profile
+\hl{Box 1: Learner Profile}
 
-Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
+\hl{Jorge has just moved from Costa Rica to Canada to study agricultural engineering.
 While fluent in both Spanish and English,
@@
 Jorge's most significant previous experience with computers is
-helping his sister build a WordPress site for the family business.
+helping his sister build a WordPress site for the family business.}
 
-Jorge needs to measure properties of soil from nearby farms
+\hl{Jorge needs to measure properties of soil from nearby farms
 using a handheld device that sends text files to his computer.
@@
 select the right values from each file,
-and calculate the required statistics.
+and calculate the required statistics.}
 
@@
 \bibitem{proghist}
-Afanador-Llach MJ, Baker J, Crymble A, Gayol V, Gibbs F, Lincoln M, McDaniel C, Milligan I, Parr J, Castro AR, Knuppel AS, Visconti A, Walsh B, Wieringa J.
-\newblock Programming Historian,
+\hl{Afanador-Llach MJ, Baker J, Crymble A, Gayol V, Gibbs F, Lincoln M, McDaniel C, Milligan I, Parr J, Castro AR, Knuppel AS, Visconti A, Walsh B, Wieringa J.}
+\newblock \hl{Programming Historian,}
 \newblock \url{https://programminghistorian.org/}.
